---
title: "Running a Simple Model & Finding a Routine"
author: "Norwegian University of Science and Technology - A Course in `MplusAutomation`"
subtitle: '*Adam Garber*'
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
library(here);library(glue)
```

______________________________________________

## Guide
   
1.  create an R project in a dedicated project folder (on the desktop or in a designated project folder)
2.  install & load packages
3.  read in data to R
4.  view data in R
5.  view metadata (from SPSS files)
6.  write .sav / .csv / .dat files
7.  fix character names to have less than 8 character 
8. introduction to mplusObjects

______________________________________________ 

## Install the “rhdf5” package to read `gh5` files (plot information)
```{r,eval=FALSE}

if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
 BiocManager::install("rhdf5")
 
```

## Load packages
```{r,eval=TRUE}
library(tidyverse)
library(haven)
library(here)
library(MplusAutomation)
```

______________________________________________

## Read in data 

______________________________________________


```{r}

# object_name <- function_1("dataset_name.sav")

exp_data <- read_spss("https://garberadamc.github.io/project-site/data/explore_lab_data.sav")

```

______________________________________________

## View dataframe with labels & response scale meta-data 

Note: Use the "print" option to save a PDF as a codebook of metadata.

______________________________________________

```{r, eval=FALSE}
# the {haven} package keeps the meta-data from SPSS files

# package_name::function_within_package()

sjPlot::view_df(exp_data)

```

______________________________________________

## Types of data for different tasks

- `.sav` (e.g., `spss_data.sav`): this data format is for SPSS files & contains variable labels (meta-data)
- `.csv` (e.g., `r_ready_data.csv`): this is the preferable data format for reading into `R` & when estimating models with `MplusAutomation` (non-labeled data)
- `.dat` (e.g., `mplus_data.dat`): this is the data format used to read into Mplus (no column names or strings)

NOTE: Mplus also accepts `.txt` formatted data (e.g., `mplus_data.txt`)

______________________________________________

## Writing, reading, and converting data between 3 formats 

______________________________________________

## Location, location, location!

NOTE: default directory in an `Rproject` is the "top-most" project folder
```{r}

here()

```


Prepare datasets, **remove SPSS labeling**
```{r,eval=FALSE}

# write a CSV datafile (preferable format for reading into R; !! removes labels !!)
write_csv(exp_data, here("02-run-models", "data", "exp_lab1_data.csv"))

# write a SPSS datafile (preferable format for reading into SPSS, labels are preserved)
write_sav(exp_data, here("02-run-models", "data", "exp_lab1_data.sav"))

```

```{r}
# read the unlabeled data back into R
nolabel_data <- read_csv(here("02-run-models", "data", "exp_lab1_data.csv"))
```

```{r,eval=FALSE}

# write a DAT datafile (this function removes header row & converts missing values to non-string)
prepareMplusData(nolabel_data, here("02-run-models", "data", "exp_lab1_data.dat"))

```

## THINGS TO NOTE:

1. This function produces a minimal template of input syntax for an Mplus input file. 
2. Behind the scenes `mplusObject()` will use a similar function to produce an input file & `.dat` file from the `R data.frame` that the function is given.
3. By default missing values in your R object (`NA`) are converted to `.` 
4. Note that if the following `file.string` is greater than **90 characters** (Mplus limit) your model will not estimate and return an error message.

  `DATA: FILE = "/Users/agarber/github/NTNU-workshop/02-run-models/data/exp_lab1_data.dat";`

______________________________________________

## Preparing column-names to be `MplusAutomation` ready

Task: Make all variable names fit within the 8-character name limit (`Mplus`) while avoiding duplicates.

______________________________________________

## Renaming columns manually...
```{r}
# use function: rename(new_name = old_name)
new_names <- nolabel_data %>% 
  rename( school_motiv1 = item1 ,  
          school_motiv2 = item2 ,
          school_motiv3 = item3 ,
          school_comp1  = item4 ,
          school_comp2  = item5 ,
          school_comp3  = item6 ,
          school_belif1 = item7 ,
          school_belif2 = item8 ,
          school_belif3 = item9 )

```

## What do you do if you have a large dataset with many column names that are > 8 characters?

- first, remove all characters greater than 8 using str_sub()
- second, make sure you don't now have duplicate variable names
- third, locate and change all duplicate names

NOTE: Use regular expressions to remove strings in a more systematic manner to avoid duplicates.

```{r}

# remove characters from the variable names that are greater than 8 characters 
names(new_names) <- str_sub(names(new_names), 1, 8)

# check if culumn names are unique 
test.unique <- function(df) {  ## function to identify unique columns
  
  length1 <- length(colnames(df))
  length2 <- length(unique(colnames(df)))        
  if (length1 - length2 > 0 ) {
    print(paste("There are", length1 - length2, " duplicates", sep=" "))
  }     
}

test.unique(new_names)

# locate duplicates (this will find the first duplicated column)
anyDuplicated(colnames(new_names))

```

______________________________________________

## A minimal example of writing, running, & reading models

______________________________________________

**What does the `mplusObject()` function do:**

1. It generates an Mplus **input** file (does not need full variable name list, its automated for you!)
2. It generates a **data file (`.dat`)** specific to each model 
3. It **runs** or estimates the model (hopefully) producing the correct output. **Always check!**

NOTE: Within the `mplusObject()` function there is a mix of `R` and `Mplus` syntax. In general the black colored text will be R code and the green colored text within the quotation marks will be Mplus code (exception; when using {`glue`}). 

______________________________________________

## PRACTICE: Using MplusObject() method 

Model is `type = BASIC;` (i.e., returns descrptive stats)
```{r,eval=FALSE}

m_basic  <- mplusObject(
  TITLE = "PRACTICE 01 - Explore TYPE = BASIC", 
  VARIABLE = 
 "usevar=
  item1 item2 item3 item4 item5
  item6 item7 item8 item9 female; 
  
  ! use exclamation symbol to make comments, reminders, or annotations in Mplus files ",
  
  ANALYSIS = 
 "type = basic; ",
 
  usevariables = colnames(nolabel_data), 
  rdata = nolabel_data)

m_basic_fit <- mplusModeler(m_basic, 
               dataout=here("02-run-models", "basic_mplus", "basic.dat"),
               modelout=here("02-run-models", "basic_mplus", "basic.inp"),
               check=TRUE, run = TRUE, hashfilename = FALSE)

```

## Check your model (Always!)

- In the `bottom-right` pane under the `files` tab click on the `basic_mplus` folder
- There should be 3 new files in this location that were produced by `mplusObject()`
- Click on the `.out` file to check if the model estimated or if there are any error messages
______________________________________________

## PRACTICE SUBSETTING: Now explore descriptives for observations that reported as "female" 

Add line of syntax: `useobs = female == 1;`


```{r,eval=FALSE}

fem_basic  <- mplusObject(
  TITLE = "PRACTICE 02 - Explore female observations only", 
  VARIABLE = 
 "usevar=
  item1 item2 item3 item4 item5
  item6 item7 item8 item9; 
  
  useobs = female == 1; !include observations that report female in analysis",
  
  ANALYSIS = 
 "type = basic;",
 
  usevariables = colnames(nolabel_data), 
  rdata = nolabel_data)

fem_basic_fit <- mplusModeler(fem_basic, 
                 dataout=here("02-run-models", "basic_mplus", "fem_basic.dat"),
                 modelout=here("02-run-models", "basic_mplus", "fem_basic.inp"),
                 check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

## PRACTICE: Exploratory Factor Analysis (EFA) 

______________________________________________

```{r,eval=FALSE}

## EXPLORATORY FACTOR ANALYSIS LAB DEMONSTRATION

efa_demo  <- mplusObject(
  TITLE = "EXPLORATORY FACTOR ANALYSIS - LAB DEMO", 
  VARIABLE = 
 "usevar=
  item1 item2 item3 item4 item5
  item6 item7 item8 item9;" ,
  
  ANALYSIS = 
 "type = efa 1 5;
  estimator = MLR;
  parallel=50;",
  
  MODEL = "" ,
  
  PLOT = "type = plot3;",
 
  OUTPUT = "sampstat standardized residual modindices (3.84);",
 
  usevariables = colnames(nolabel_data), 
  rdata = nolabel_data)

efa_demo_fit <- mplusModeler(efa_demo, 
  dataout=here("02-run-models", "basic_mplus", "EFA_Lab_DEMO.dat"),
  modelout=here("02-run-models", "basic_mplus", "EFA_Lab_DEMO.inp"),
  check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

**END**

______________________________________________











