---
title: "Splits & Iterators"
author: "Norwegian University of Science and Technology - A Course in `MplusAutomation`"
subtitle: '*Adam Garber*'
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
library(here);library(glue)
```

______________________________________________

### Outline

1. Randomly split data into 2 equal parts (calibration & validation samples)
2. Introduction to using iterators or loops with `mplusObject()`

____________________________________

### Getting started - following the routine...

1. Create an R-Project
2. Install packages ($\color{red}{\text{ONLY IF NEEDED}}$)
3. Load packages

### Folder structure:

Parent folder:
`03-efa`

Nested folders:

- `data`
- `efa_mplus`
- `figures`

____________________________________

New packages:

- {`janitor`}

____________________________________

## Begin 

____________________________________

DATA SOURCE: This lab exercise utilizes the NCES public-use dataset: Education Longitudinal Study of 2002 (Lauff & Ingels, 2014) [$\color{blue}{\text{See website: nces.ed.gov}}$](https://nces.ed.gov/surveys/els2002/avail_data.asp)

____________________________________

### loading packages...

```{r, eval=TRUE}
library(janitor)
library(tidyverse)
library(haven)
library(MplusAutomation)
library(here)
library(corrplot)
library(glue)
```

## Change starting location to folder `04-splits-iterators`
```{r}

source("rep_functions.R")

change_here(glue("{project_location}/04-splits-iterators"))

here()
```

### read in the raw dataset 

```{r}
lab_data <- read_spss("https://garberadamc.github.io/project-site/data/els_sub1_spss.sav")
```

### create a subset of the dataset called `school_trouble`
```{r}
school_trouble <- lab_data %>% 
  select(41:55)
```

### make a new codebook from the `school_trouble` subset
```{r, eval=FALSE}
sjPlot::view_df(school_trouble)
```

### write a `CSV` datafile
```{r, eval=FALSE}
write_csv(school_trouble, here("data", "school_trouble_data.csv"))
```

### read the unlabeled data back into R
```{r}
trouble_data <- read_csv(here("data", "school_trouble_data.csv"))
```

### check items to see if reverse coding is needed 
```{r}
cor_matrix <- cor(trouble_data, use = "pairwise.complete.obs")

corrplot(cor_matrix, method="circle",
         type = "upper",
         tl.col="black", 
         tl.srt=45)
```

____________________________________

## Randomly split a sample into 2 equal parts

____________________________________

- Get *n*-size of half of original sample using `nrow()`
- The `floor()` function helps with rounding
```{r}
smp_size <- floor(0.50 * nrow(trouble_data))
```

### set the seed to make your partition reproducible
```{r}
set.seed(123)
```

### the function `sample()` will pick at random the values of the specified number
```{r}
calibrate_smp <- sample(seq_len(nrow(trouble_data)), size = smp_size)
```

### create two samples called "calibrate" & "validate"
```{r}
calibrate <- trouble_data[calibrate_smp, ]
validate <- trouble_data[-calibrate_smp, ]
```

### Run EFA with the `calibrate` sample
```{r, eval=FALSE}
m_efa_1  <- mplusObject(
  TITLE = "School Trouble EFA - LAB 4 DEMO", 
  VARIABLE = 
    "usevar = BYS22A-BYS24G;", 
  
  ANALYSIS = 
    "type = efa 1 5;   
     estimator = mlr;  
     parallel=50; ! run parallel analysis",
  
  MODEL = "" ,
  
  PLOT = "type = plot3;",
  OUTPUT = "sampstat;",
  
  usevariables = colnames(calibrate), 
  rdata = calibrate)

m_efa_1_fit <- mplusModeler(m_efa_1, 
               dataout=here("efa_mplus", "efa1_trouble.dat"),
               modelout=here("efa_mplus", "efa1_trouble.inp"),
               check=TRUE, run = TRUE, hashfilename = FALSE)
```

### Plot Parallel Analysis & Eigenvalues

### read into R an Mplus output file
```{r}
efa_summary <- readModels(here("efa_mplus", "efa1_trouble.out"), quiet = TRUE)
``` 

### extract relavent data & prepare dataframe for plot
```{r}
plot_data <- list(EFA=efa_summary[["gh5"]][["efa"]][["eigenvalues"]], 
             Parallel=efa_summary[["gh5"]][["efa"]][["parallel_average"]]) %>% 
  as.data.frame() %>% 
  mutate(Factor = row_number()) %>% 
  mutate(Factor = factor(Factor))
```

### pivot the dataframe to "long" format
```{r}
plot_data_long <- plot_data %>% 
  pivot_longer(EFA:Parallel,               # The columns I'm gathering together
               names_to = "Analysis",      # new column name for existing names
               values_to = "Eigenvalues")  # new column name to store values
``` 

### plot using ggplot
```{r}
plot_data_long %>% 
  ggplot(aes(y=Eigenvalues,
             x=Factor,
             group=Analysis,
             color=Analysis)) +
  geom_point() + 
  geom_line() + 
  theme_minimal()
``` 

### save figure to the designated folder
```{r, eval=FALSE}
ggsave(here("figures", "eigenvalue_elbow_rplot.png"), dpi=300, height=5, width=7, units="in")
```

____________________________________

## Introduction to MplusAutomation with iterators 

____________________________________

### Alternate way to run an EFA with the "calibrate" sample
```{r, eval=FALSE}
m_efa_k15  <- lapply(1:5, function(k) {
  
  m_efa  <- mplusObject(
    TITLE = "School Trouble EFA - LAB 4 DEMO", 
    VARIABLE = 
      "usevar = BYS22A-BYS24G;", 
    
    ANALYSIS = 
      paste("type=efa", k, k), 
    
    MODEL = "" ,
    
    PLOT = "type = plot3;",
    OUTPUT = "sampstat;",
    
    usevariables = colnames(calibrate), 
    rdata = calibrate)
  
  m_efa_fit <- mplusModeler(m_efa, 
               dataout=sprintf(here("efa_mplus2", "efa_trouble.dat"), k),
               modelout=sprintf(here("efa_mplus2", "efa_%d_trouble.inp"), k),
               check=TRUE, run = TRUE, hashfilename = FALSE)
  
})
```


## References

Hallquist, M. N., & Wiley, J. F. (2018). MplusAutomation: An R Package for Facilitating Large-Scale Latent Variable Analyses in Mplus. Structural equation modeling: a multidisciplinary journal, 25(4), 621-638.

Horst, A. (2020). Course & Workshop Materials. GitHub Repositories, https://https://allisonhorst.github.io/

Muthén, L.K. and Muthén, B.O. (1998-2017).  Mplus User’s Guide.  Eighth Edition. Los Angeles, CA: Muthén & Muthén

R Core Team (2017). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL http://www.R-project.org/

Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686
