---
title: "Confirmatory Factor Analysis (CFA) - School Trouble Example"
author: "Norwegian University of Science and Technology - A Course in `MplusAutomation`"
subtitle: '*Adam Garber*'
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
library(here);library(glue)
```

______________________________________________

## Outline

1. One more quick EFA: change rotation method (override the default)
2. Specify a confirmatory factor analysis (CFA)
3. Generate a path diagram of your CFA model

______________________________________________

## Getting started - following the routine...

1. Create an R-Project
2. Install packages 
3. Load packages

### R-Project instructions:

1. click "NEW PROJECT" (upper right corner of window)
2. choose option "NEW DIRECTORY"
3. choose location of project (on desktop OR in a designated folder)

Within R-studio under the files pane (bottom right):

1. click "New Folder" and name folder "data"
2. click "New Folder" and name folder "efa_mplus"
3. click "New Folder" and name folder "cfa_mplus"

______________________________________________

DATA SOURCE: This lab exercise utilizes the NCES public-use dataset: Education Longitudinal Study of 2002 (Lauff & Ingels, 2014) [$\color{blue}{\text{See website: nces.ed.gov}}$](https://nces.ed.gov/surveys/els2002/avail_data.asp)

______________________________________________

## Begin 

______________________________________________

loading packages...
```{r, eval=TRUE}
library(tidyverse)
library(MplusAutomation)
library(here)
library(semPlot) 
library(stargazer)
```


read in data
```{r, eval=TRUE}

trouble_data <- read_csv("https://garberadamc.github.io/project-site/data/els_sub3_school_trouble.csv")

```

take a look at the EFA data (same indicators used for lab 4)
```{r}

stargazer(as.data.frame(trouble_data), type="text", digits=1)

```

replicate the SAME random split as last week (do not change the seed)

- get the size of half of original sample
- `floor()` helps with rounding

```{r}
smp_size <- floor(0.50 * nrow(trouble_data))
```

set the seed to make partition reproducible
```{r}
set.seed(20210530)
```

`sample()` randomly selects rows of size = "smp_size"
```{r}
calibrate_smp <- sample(seq_len(nrow(trouble_data)), size = smp_size)
```

create two samples called "calibrate" & "validate"
```{r}
calibrate <- trouble_data[calibrate_smp, ]
validate <- trouble_data[-calibrate_smp, ]
```

reorder variables & remove column number 15 
```{r}

# removed BYS24G (due to low loading < .2)
calibrate <-  calibrate %>% 
  select(1,3:14,2) 

# do the same for the validate sample
validate <-  validate %>% 
  select(1,3:14,2) 
```


### Default rotation: Geomin Oblique

run an EFA with the "calibrate" sample
```{r, eval = FALSE}
efa_geomin  <- mplusObject(
  TITLE = "Geomin Oblique EFA - LAB 5 DEMO", 
  VARIABLE = 
   "! removed BYS24G (due to low loading < .2)
    usevar = BYS22A-BYS22B;", 
  
  ANALYSIS = 
   "type = efa 1 3;   
    estimator = mlr;
    rotation = geomin; ! this is the default (added to be explicit)
    parallel=50; ! run parallel analysis",
  
  MODEL = "" ,
  
  PLOT = "type = plot3;",
  OUTPUT = "sampstat;",
  
  usevariables = colnames(calibrate), 
  rdata = calibrate)

efa_geomin_fit <- mplusModeler(efa_geomin, 
                  dataout=here("05-cfa", "efa_mplus", "efa_geomin_oblique.dat"),
                  modelout=here("05-cfa", "efa_mplus", "efa_geomin_oblique.inp"),
                  check=TRUE, run = TRUE, hashfilename = FALSE)
```

### Rotation: Varimax Orthogonal
```{r, eval = FALSE}
efa_varimax  <- mplusObject(
  TITLE = "Varimax Orthogonal EFA - LAB 5 DEMO", 
  VARIABLE = 
    " ! removed BYS24G (due to low loading < .2)
    usevar = BYS22A-BYS22B;", 
  
  ANALYSIS = 
    "type = efa 1 3;   
     estimator = mlr;
     rotation = varimax; ! orthogonal (no factor correlations)
     parallel=50; ! run parallel analysis",
  
  MODEL = "" ,
  
  PLOT = "type = plot3;",
  OUTPUT = "sampstat;",
  
  usevariables = colnames(calibrate), 
  rdata = calibrate)

efa_varimax_fit <- mplusModeler(efa_varimax, 
                            dataout=here("05-cfa", "efa_mplus", "efa_varimax_orthogonal.dat"),
                            modelout=here("05-cfa", "efa_mplus", "efa_varimax_orthogonal.inp"),
                            check=TRUE, run = TRUE, hashfilename = FALSE)
```

### Estimate a Confirmatory Factor Analysis (CFA) model
```{r, eval = FALSE}
cfa_validate  <- mplusObject(
  
  TITLE = "CFA Validate", 
  
  VARIABLE = 
    "usevar = BYS22A-BYS22B;", 
  
  ANALYSIS = 
    "estimator = mlr;",
  
  MODEL = 
    "FACTOR_1 by BYS22A BYS22C BYS22D BYS22E BYS22F BYS22G BYS22H;
  
     FACTOR_2 BY BYS24A BYS24B BYS24C BYS24D BYS24E BYS24F BYS22B;" ,
  
  PLOT = "type = plot3;",
  
  OUTPUT = "sampstat standardized residual modindices (3.84);",
  
  usevariables = colnames(validate), 
  rdata = validate)

cfa_val_fit <- mplusModeler(cfa_validate, 
                            dataout=here("05-cfa", "cfa_mplus", "lab5_cfa_validate.dat"),
                            modelout=here("05-cfa", "cfa_mplus", "lab5_cfa_validate.inp"),
                            check=TRUE, run = TRUE, hashfilename = FALSE)
```

#### create a path diagram of the CFA model

Read in the model to R within the `cfa_mplus` folder
```{r}
cfa_output <- readModels(here("05-cfa", "cfa_mplus"), quiet = TRUE)
```

Plot model
```{r}
semPaths(cfa_output,
         intercepts=FALSE,
         fixedStyle = c(1)
         )
```

play with some formatting
```{r}
semPaths(cfa_output, "std",
         intercepts=FALSE,
         fixedStyle = c(1),
         fade = FALSE,
         color= list(lat = c("light blue"," light green")),
        )

```

______________________________________________

### End 

______________________________________________

## References

Hallquist, M. N., & Wiley, J. F. (2018). MplusAutomation: An R Package for Facilitating Large-Scale Latent Variable Analyses in Mplus. Structural equation modeling: a multidisciplinary journal, 25(4), 621-638.

Horst, A. (2020). Course & Workshop Materials. GitHub Repositories, https://https://allisonhorst.github.io/

Muthén, L.K. and Muthén, B.O. (1998-2017).  Mplus User’s Guide.  Eighth Edition. Los Angeles, CA: Muthén & Muthén

R Core Team (2017). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL http://www.R-project.org/

Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686
