# Principle Component Analysis (PCA)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, message = FALSE, warning = FALSE)
```

____________________________________

### Outline Chapter 7 - Principle Component Analysis:

1. 10th grade student demographics & school safety (ELS, 2002 public-use data)

**NOTE:** Syntax is modeled after Allison Horst's PCA lab at UCSB (ESM-206; Horst, 2020) using an example applied to Education data.

____________________________________

DATA SOURCE: This lab exercise utilizes the NCES public-use dataset: Education Longitudinal Study of 2002 (Lauff & Ingels, 2014) [$\color{blue}{\text{See website: nces.ed.gov}}$](https://nces.ed.gov/surveys/els2002/avail_data.asp)

____________________________________

### load packages
```{r, eval=TRUE}
library(FactoMineR)
library(factoextra)
library(skimr)
library(naniar)
library(ggfortify)
library(janitor)
library(tidyverse)
library(here)
```

### read in ELS-2002 lab data:
```{r}
lab_data <- read_csv("https://garberadamc.github.io/project-site/data/els_sub4.csv")
```

### make all column names "lower_snake_case" style
```{r}
lab_tidy <- lab_data %>% 
  clean_names()
```

### Prepare data for PCA
```{r}
# remove variables that don't make sense in a PCA
lab_sub1 <- lab_tidy %>% 
  select(-stu_id,   # these are random numbers
         -sch_id,
         -byrace,   # nominal (non-ordered variable) 
         -byparace, # nominal (non-ordered variable) 
         -byparlng, # nominal (non-ordered variable)
         -byfcomp,  # nominal (non-ordered variable)
         -bypared, -bymothed, -byfathed,
         -bysctrl, -byurban, -byregion)

# select columns and rename variables to have descriptive names
lab_sub2 <- lab_sub1 %>% 
  select(1:9,
         bys20a, bys20h, bys20j, bys20k, bys20m, bys20n,
         bys21b, bys21d, bys22a, bys22b, bys22c, bys22d,
         bys22e, bys22g, bys22h, bys24a, bys24b) %>% 
  rename("stu_exp" = "bystexp",
         "par_asp" = "byparasp",
         "mth_read" = "bytxcstd",
         "mth_test" = "bytxmstd",
         "rd_test" = "bytxrstd",
         "freelnch" = "by10flp",
         "stu_tch" = "bys20a",
         "putdownt" = "bys20h",
         "safe" = "bys20j",
         "disrupt" = "bys20k",
         "gangs" = "bys20m",
         "rac_fght" = "bys20n",
         "fair" = "bys21b",
         "strict" = "bys21d",
         "stolen" = "bys22a",
         "drugs" = "bys22b",
         "t_hurt" = "bys22c",
         "p_fight" = "bys22d",
         "hit" = "bys22e",
         "damaged" = "bys22g",
         "bullied" = "bys22h",
         "late" = "bys24a",
         "skipped" = "bys24b")

```

### Investigate missingness {`naniar`} & make data summary with {`skimr`}
```{r}
# Plot number of missings by variable
gg_miss_var(lab_sub2)

# Look at summary of data using skimr::skim()
skim(lab_sub2)

pca1 <- lab_sub2 %>% 
  drop_na()
```

### run PCA with `prcomp()` (function does not permit NA values)
```{r, eval = FALSE}

pca_out1 <- prcomp(pca1, scale = TRUE) 

plot(pca_out1)

#summary(pca_out1)
```

### plot PCA biplot
```{r}

jpeg(here("figures", "biplot_pca1.jpg"), res = 100) # to save the biplot

my_biplot <- autoplot(pca_out1, 
                      colour = NA,
                      loadings.label = TRUE,
                      loadings.label.size = 3,
                      loadings.label.colour = "black",
                      loadings.label.repel = TRUE) +
            theme_minimal()
  
my_biplot

dev.off()
```

```{r}
my_biplot
```

### alternative funtion to run & plot PCA biplot
```{r}
PCA(pca1, scale.unit = TRUE, ncp = 20, graph = TRUE)
```

## References

Hallquist, M. N., & Wiley, J. F. (2018). MplusAutomation: An R Package for Facilitating Large-Scale Latent Variable Analyses in Mplus. Structural equation modeling: a multidisciplinary journal, 25(4), 621-638.

Horst, A. (2020). Course & Workshop Materials. GitHub Repositories, https://https://allisonhorst.github.io/

Muthén, L.K. and Muthén, B.O. (1998-2017).  Mplus User’s Guide.  Eighth Edition. Los Angeles, CA: Muthén & Muthén

R Core Team (2017). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL http://www.R-project.org/

Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686