---
title: "Measurement Invariance"
author: "Norwegian University of Science and Technology - A Course in `MplusAutomation`"
subtitle: '*Adam Garber*'
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
library(here);library(glue);library(linguisticsdown)
```

______________________________________________

DATA SOURCE: This lab exercise utilizes the NCES public-use dataset: Education Longitudinal Study of 2002 (Lauff & Ingels, 2014) [$\color{blue}{\text{See website: nces.ed.gov}}$](https://nces.ed.gov/surveys/els2002/avail_data.asp)

```{r, eval=TRUE}
# load packages
library(MplusAutomation)
library(tidyverse)
library(here)
library(corrplot)
library(gt)
library(reshape2)
library(semPlot)
```

## Change starting location to folder `09-invariance`
```{r}

source("rep_functions.R")

change_here(glue("{project_location}/09-invariance"))

here()
```

______________________________________________

## Lab 9 - Begin 

______________________________________________


Read in data
```{r}

lab_data <- read_csv("https://garberadamc.github.io/project-site/data/els_sub5_data.csv")

```

Preparations: subset, reorder, rename, and recode data
```{r}

invar_data <-  lab_data %>% 
  select(bystlang, freelnch, byincome,                    # covariates 
         stolen, t_hurt, p_fight, hit, damaged, bullied,  # factor 1 (indicators)
         safe, disrupt, gangs, rac_fght,                  # factor 2 (indicators)
         late, skipped, mth_read, mth_test, rd_test) %>%  
  rename("unsafe" = "safe") %>% 
  mutate(freelnch = case_when( # Grade 10, percent free lunch - transform to binary
        freelnch <  3 ~ 0,     # school has less than 11% 
        freelnch >= 3 ~ 1))    # school has greater than or equal to 11% 

table(invar_data$freelnch) # reasonably balanced groups

```

Take a quick look at variable distributions
```{r}

melt(invar_data[,4:13]) %>% 
  ggplot(., aes(x=value, label=variable)) +
  geom_histogram(bins = 15) +
  facet_wrap(~variable, scales = "free")

```

Reverse code factor for ease of interpretation 
```{r}

cols = c("unsafe", "disrupt", "gangs", "rac_fght")

invar_data[ ,cols] <-  5 - invar_data[ ,cols]

```

Factor names and interpretation:

- VICTIM: student reports being a victim of injury to self or property
    - scale range: Never, Once or twice, More than twice
    - higher values indicate greater frequency of victimization events

- NEG_CLIM: Student reports on negative school climate attributes 
    - scale range: Strongly Disagree - Strongly Agree
    - higher values indicate a more negative climate


Check correct coding, explore correlations 
```{r}

cor_matrix <- cor(invar_data[4:13], use = "pairwise.complete.obs")

corrplot(cor_matrix, 
         method = "circle",
         type = "upper")
```

______________________________________________

## Estimate the Unconditional Confirmatory Factor Analysis (CFA) model

______________________________________________

Number of parameters = 31

- 10 item loadings
- 10 intercepts
- 10 residual variances
- 01 factor co-variances 

```{r, eval=FALSE}
cfa_m0  <- mplusObject(
  TITLE = "model0 - unconditional CFA model", 
  VARIABLE = 
    "usevar = stolen-rac_fght;", 
  
  ANALYSIS = 
    "estimator = mlr;",
  
  MODEL = 
    "VICTIM by stolen* t_hurt p_fight hit damaged bullied;
     VICTIM@1; ! UVI identification
     
     NEG_CLIM by unsafe* disrupt gangs rac_fght;
     NEG_CLIM@1; ",
  
  PLOT = "type = plot3;",
  OUTPUT = "sampstat standardized residual modindices (3.84);",
  
  usevariables = colnames(invar_data), 
  rdata = invar_data)

cfa_m0_fit <- mplusModeler(cfa_m0, 
              dataout=here("invar_mplus", "invar_data.dat"),
              modelout=here("invar_mplus", "m0_cfa_fullsample.inp"),
              check=TRUE, run = TRUE, hashfilename = FALSE)
```


______________________________________________

## Run separate CFA models for each sub-sample

______________________________________________

### Group `freelnch = 0` (low) CFA
```{r, eval=FALSE}

cfa_m1  <- mplusObject(
  TITLE = "CFA model1 - group is 0 for freelnch", 
  VARIABLE = 
    "usevar = stolen-rac_fght;
     
     !freelnch (0 = school proportion is less than 11 percent)
     USEOBS = freelnch == 0; ", 
  
  ANALYSIS = 
    "estimator = mlr;",
  
  MODEL = 
    "VICTIM by stolen* t_hurt p_fight hit damaged bullied;
     VICTIM@1; ! UVI identification
     
     NEG_CLIM by unsafe* disrupt gangs rac_fght;
     NEG_CLIM@1; ",
  
  PLOT = "type = plot3;",
  OUTPUT = "sampstat standardized residual modindices (3.84);",
  
  usevariables = colnames(invar_data), 
  rdata = invar_data)

cfa_m1_fit <- mplusModeler(cfa_m1, 
              dataout=here("invar_mplus", "invar_data.dat"),
              modelout=here("invar_mplus", "m1_cfa_freelnch_0.inp"),
              check=TRUE, run = TRUE, hashfilename = FALSE)
```

### Group `freelnch = 1` (moderate to high)  CFA
```{r, eval=FALSE}

cfa_m2  <- mplusObject(
  
  TITLE = "CFA model2 - group is 1 for freelnch", 
  
  VARIABLE = 
    "usevar = stolen-rac_fght;
     
     !freelnch (1 = school proportion is greater than or equal to 11 percent)
     USEOBS = freelnch == 1; ", 
  
  ANALYSIS = 
    "estimator = mlr;",
  
  MODEL = 
    "VICTIM by stolen* t_hurt p_fight hit damaged bullied;
     VICTIM@1; ! UVI identification
     
     NEG_CLIM by unsafe* disrupt gangs rac_fght;
     NEG_CLIM@1; ",
     
  PLOT = "type = plot3;",
  
  OUTPUT = "sampstat standardized residual modindices (3.84);",
  
  usevariables = colnames(invar_data), 
  rdata = invar_data)

cfa_m2_fit <- mplusModeler(cfa_m2, 
              dataout=here("invar_mplus", "invar_data.dat"),
              modelout=here("invar_mplus", "m2_cfa_freelnch_1.inp"),
              check=TRUE, run = TRUE, hashfilename = FALSE)

```

## Multi-Group Invariance Models

```{r, echo=FALSE, eval=TRUE, out.width = "70%", out.height= "70%", fig.align = "center"}
include_graphics2("https://garberadamc.github.io/project-site/figures/mean_structure.png")
```
*Figure*: Picture depicting mean structure from slide by Dr. Karen Nylund-Gibson

______________________________________________

## Configural invariance

- free item loadings, intercepts, and residuals
- factor means fixed to zero
- factor variances fixed to 1

______________________________________________

Number of parameters = 62

- 20 item loadings (10items*2groups)
- 20 intercepts
- 20 residual variances
- 02 factor co-variances (1 for each group)

______________________________________________

```{r, eval=FALSE}

cfa_m3  <- mplusObject(
  
  TITLE = "CFA model3 - configural invariance", 
  
  VARIABLE = 
    "usevar = stolen-rac_fght;
    
     grouping = freelnch (0=freelnch_0 1=freelnch_1); ", 
  
  ANALYSIS = 
    "estimator = mlr;",
  
  MODEL = 
    "VICTIM by stolen* t_hurt p_fight hit damaged bullied;
     VICTIM@1; ! UVI identification
     
     NEG_CLIM by unsafe* disrupt gangs rac_fght;
     NEG_CLIM@1;
     
     [VICTIM-NEG_CLIM@0]; !factor means set to zero
  
     MODEL freelnch_1:
     
     VICTIM by stolen* t_hurt p_fight hit damaged bullied;
     VICTIM@1; 
     
     [stolen t_hurt p_fight hit damaged bullied]; !free intercepts
     
     NEG_CLIM by unsafe* disrupt gangs rac_fght;
     NEG_CLIM@1;
     
     [unsafe disrupt gangs rac_fght]; !free intercepts
     
     [VICTIM-NEG_CLIM@0]; ",
  
  PLOT = "type = plot3;",
  OUTPUT = "sampstat standardized residual modindices (3.84);",
  
  usevariables = colnames(invar_data), 
  rdata = invar_data)

cfa_m3_fit <- mplusModeler(cfa_m3, 
              dataout=here("invar_mplus", "invar_data.dat"),
              modelout=here("invar_mplus", "m3_configural.inp"),
              check=TRUE, run = TRUE, hashfilename = FALSE)

```


______________________________________________

## Metric invariance

- item loadings (set to equal)
- free intercepts and residuals
- factor means fixed to zero
- free factor variances in group 2

______________________________________________

Number of parameters = 54

- 10 item loadings (set to equal)
- 20 intercepts
- 20 residual variances
- 02 factor variances 
- 02 factor co-variances 

______________________________________________

```{r, eval=FALSE}

cfa_m4  <- mplusObject(
  TITLE = "CFA model4 - metric invariance", 
  VARIABLE = 
    "usevar = stolen-rac_fght;
    
     grouping = freelnch (0=freelnch_0 1=freelnch_1); ", 
  
  ANALYSIS = 
    "estimator = mlr;",
  
  MODEL = 
    "VICTIM by stolen* t_hurt p_fight hit damaged bullied;
     VICTIM@1; ! UVI identification
     
     NEG_CLIM by unsafe* disrupt gangs rac_fght;
     NEG_CLIM@1;
     
     [VICTIM-NEG_CLIM@0]; 
  
     MODEL freelnch_1:
     
     VICTIM; ! free factor variances for group 2
     
     [stolen t_hurt p_fight hit damaged bullied]; 
     
     NEG_CLIM;
     
     [unsafe disrupt gangs rac_fght]; 
     
     [VICTIM-NEG_CLIM@0]; ",
  
  PLOT = "type = plot3;",
  OUTPUT = "sampstat standardized residual modindices (3.84);",
  
  usevariables = colnames(invar_data), 
  rdata = invar_data)

cfa_m4_fit <- mplusModeler(cfa_m4, 
              dataout=here("invar_mplus", "invar_data.dat"),
              modelout=here("invar_mplus", "m4_metric.inp"),
              check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

## Scalar invariance

- item loadings (set to equal)
- intercepts (set to equal)
- free residuals
- free factor variances and means in group 2

______________________________________________

Number of parameters = 46

- 10 item loadings (set to equal)
- 10 intercepts (set to equal)
- 20 residual variances
- 02 factor variances 
- 02 factor co-variances 
- 02 factor means

______________________________________________

```{r, eval=FALSE}

cfa_m5  <- mplusObject(
  TITLE = "model5 - scalar invariance", 
  VARIABLE = 
    "usevar = stolen-rac_fght;
    
     grouping = freelnch (0=freelnch_0 1=freelnch_1); ", 
  
  ANALYSIS = 
    "estimator = mlr;",
  
  MODEL = 
    "VICTIM by stolen* t_hurt p_fight hit damaged bullied;
     VICTIM@1; 
     
     NEG_CLIM by unsafe* disrupt gangs rac_fght;
     NEG_CLIM@1;
     
     [VICTIM-NEG_CLIM@0]; 
  
     MODEL freelnch_1:
     
     VICTIM; ! free factor variances for group 2
     
     NEG_CLIM;
     
     [VICTIM-NEG_CLIM]; ! free factor means", 
  
  PLOT = "type = plot3;",
  OUTPUT = "sampstat standardized residual modindices (3.84);",
  
  usevariables = colnames(invar_data), 
  rdata = invar_data)

cfa_m5_fit <- mplusModeler(cfa_m5, 
              dataout=here("invar_mplus", "invar_data.dat"),
              modelout=here("invar_mplus", "m5_scalar.inp"),
              check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

## Strict invariance

- item loadings (set to equal)
- intercepts (set to equal)
- residuals (set to equal)
- free factor variances and means in group 2

______________________________________________

Number of parameters = 36

- 10 item loadings (set to equal)
- 10 intercepts (set to equal)
- 10 residual variances
- 02 factor variances 
- 02 factor co-variances 
- 02 factor means

______________________________________________

```{r, eval=FALSE}

cfa_m6  <- mplusObject(
  TITLE = "model6 - strict invariance", 
  VARIABLE = 
    "usevar = stolen-rac_fght;
    
     grouping = freelnch (0=freelnch_0 1=freelnch_1); ", 
  
  ANALYSIS = 
    "estimator = mlr;",
  
  MODEL = 
    "VICTIM by stolen* t_hurt p_fight hit damaged bullied;
     VICTIM@1; 
     
     NEG_CLIM by unsafe* disrupt gangs rac_fght;
     NEG_CLIM@1;
     
     [VICTIM-NEG_CLIM@0]; 
     
     stolen-rac_fght(1-10); ! set residuals to be equal across groups
  
     MODEL freelnch_1:
     
     VICTIM; ! free factor variances for group 2
     
     NEG_CLIM;
     
     [VICTIM-NEG_CLIM]; ! free factor means
  
     stolen-rac_fght(1-10); ", 
  
  PLOT = "type = plot3;",
  OUTPUT = "sampstat standardized residual modindices (3.84);",
  
  usevariables = colnames(invar_data), 
  rdata = invar_data)

cfa_m6_fit <- mplusModeler(cfa_m6, 
              dataout=here("invar_mplus", "invar_data.dat"),
              modelout=here("invar_mplus", "m6_strict.inp"),
              check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

## Structural invariance A (fixed factor variances)

Demonstration of structural invariance using the **Scalar model**

- item loadings (set to equal)
- intercepts (set to equal)
- free residuals (Scalar)
- factor means free in group 2
- factor variances (set to 1)
- free factor covariances

______________________________________________

Number of parameters = 44

- 10 item loadings (set to equal)
- 10 intercepts (set to equal)
- 20 residual variances
- 00 factor variances 
- 02 factor co-variances 
- 02 factor means

______________________________________________

```{r, eval=FALSE}
# fixed factor variances

cfa_m7  <- mplusObject(
  
  TITLE = "model7 - structural invariance A" , 
  
  VARIABLE = 
    "usevar = stolen-rac_fght;
    
     grouping = freelnch (0=freelnch_0 1=freelnch_1); ", 
  
  ANALYSIS = 
    "estimator = mlr;",
  
  MODEL = 
    "VICTIM by stolen* t_hurt p_fight hit damaged bullied;
     VICTIM@1; 
     
     NEG_CLIM by unsafe* disrupt gangs rac_fght;
     NEG_CLIM@1;
     
     [VICTIM-NEG_CLIM@0]; 
  
     MODEL freelnch_1:
     
     [VICTIM-NEG_CLIM]; ! free factor means
      
     VICTIM@1; NEG_CLIM@1; ! fix factor variance to 1", 
  
  PLOT = "type = plot3;",
  OUTPUT = "sampstat standardized residual modindices (3.84);",
  
  usevariables = colnames(invar_data), 
  rdata = invar_data)

cfa_m7_fit <- mplusModeler(cfa_m7, 
              dataout=here("invar_mplus", "invar_data.dat"),
              modelout=here("invar_mplus", "m7_structuralA.inp"),
              check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

## Structural invariance B (fixed factor variances and equal covariances)

Demonstration of structural invariance using the **Scalar model**

- item loadings (set to equal)
- intercepts (set to equal)
- free residuals (Scalar)
- factor means free in group 2
- factor variances (set to equal)
- factor covariances (set to equal)

______________________________________________

Number of parameters = 43

- 10 item loadings (set to equal)
- 10 intercepts (set to equal)
- 20 residual variances
- 00 factor variances 
- 01 factor co-variances 
- 02 factor means

______________________________________________

```{r, eval=FALSE}

# equal factor variances and covariances

cfa_m8  <- mplusObject(
  TITLE = "model8 - structural invariance B" , 
  VARIABLE = 
    "usevar = stolen-rac_fght;
    
     grouping = freelnch (0=freelnch_0 1=freelnch_1); ", 
  
  ANALYSIS = 
    "estimator = mlr;",
  
  MODEL = 
    "VICTIM by stolen* t_hurt p_fight hit damaged bullied;
     VICTIM@1; 
     
     NEG_CLIM by unsafe* disrupt gangs rac_fght;
     NEG_CLIM@1;
     
     [VICTIM-NEG_CLIM@0]; 
     
     VICTIM with NEG_CLIM (11) ! set covariances to equal;
  
     MODEL freelnch_1:
     
     [VICTIM-NEG_CLIM]; ! free factor means
      
     VICTIM@1; NEG_CLIM@1; ! fix factor variance to 1
     
     VICTIM with NEG_CLIM (11); ! set covariances to equal", 
  
  PLOT = "type = plot3;",
  OUTPUT = "sampstat standardized residual modindices (3.84);",
  
  usevariables = colnames(invar_data), 
  rdata = invar_data)

cfa_m8_fit <- mplusModeler(cfa_m8, 
              dataout=here("invar_mplus", "invar_data.dat"),
              modelout=here("invar_mplus", "m8_structuralB.inp"),
              check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

## Latent Factor Means differences: 

(model: Step_07_STRUCTURAL)

Mean differences: Students in sub-sample `freelnch_1` have...

    VICTIM     -0.026      0.091     -0.291      0.771  (not significant)
    NEG_CLIM    0.632      0.104      6.104      0.000  (higher scores for "NEG_CLIM")
    

______________________________________________

## Comparing Fit Across Models

______________________________________________

## Guidlines: for loadings & fit indices

- **Simple structure:** “0.4 - 0.3 - 0.2” rule Howard (2016)
(primary loadings > 0.4 / cross-loadings < 0.3 / minimum difference = 0.2)
- **RMSEA:** < .05 indicates "good" fit Brown (2015)
-   **CFI:** > .95 indicates "good" fit Brown (2015)
-  **SRMR:** < .08 indicates "good" fit Hu and Bentler (1999)
- **Invariance:** Changes in **CFI** less than or equal to **-0.01** are acceptable

______________________________________________

Read into R summary of all models
```{r}
all_models <- readModels(here("invar_mplus"), quiet = TRUE)
```

______________________________________________

Create table, extract fit statistics, sort by `Filename`
```{r}

invar_summary <- LatexSummaryTable(all_models, 
                 keepCols=c("Filename", "Parameters","ChiSqM_Value", "CFI","TLI",
                            "SRMR", "RMSEA_Estimate", "RMSEA_90CI_LB", "RMSEA_90CI_UB"), 
                 sortBy = "Filename")

invar_summary %>%
  mutate(Filename = str_remove(Filename, ".out")) %>% 
  gt() %>% 
  cols_label(
    Filename = "Model",
    Parameters = "Par",
    ChiSqM_Value = "ChiSq",
    CFI = "CFI", TLI = "TLI", SRMR = "SRMR",
    RMSEA_Estimate = "RMSEA",
    RMSEA_90CI_LB = "Lower CI",
    RMSEA_90CI_UB = "Upper CI")
  

```

______________________________________________

## Calculate Satora-Bentler scaled Chi-square difference test (use with MLR estimator) 

[$\color{blue}{\text{See website: stats.idre.ucla.edu}}$](https://stats.idre.ucla.edu/mplus/faq/how-can-i-compute-a-chi-square-test-for-nested-models-with-the-mlr-or-mlm-estimators/)

- SB0 = null model Chi-square value
- SB1 = alternate model Chi-square value
- c0  = null model scaling correction factor
- c1  = alternate model scaling correction factor
- d0  = null model degrees of freedom
- d1  = alternate model degrees of freedom
- df  = Chi-square test degrees of freedom

______________________________________________

compare configural to metric
```{r}

SB0 <- all_models[["M4_metric.out"]][["summaries"]][["ChiSqM_Value"]]
SB1 <- all_models[["M3_configural.out"]][["summaries"]][["ChiSqM_Value"]]
c0  <- all_models[["M4_metric.out"]][["summaries"]][["ChiSqM_ScalingCorrection"]]
c1  <- all_models[["M3_configural.out"]][["summaries"]][["ChiSqM_ScalingCorrection"]]
d0  <- all_models[["M4_metric.out"]][["summaries"]][["ChiSqM_DF"]]
d1  <- all_models[["M3_configural.out"]][["summaries"]][["ChiSqM_DF"]]
df  <- abs(d0-d1)

# Satora-Bentler scaled Difference test equations
cd <- (((d0*c0)-(d1*c1))/(d0-d1))
t  <- (((SB0*c0)-(SB1*c1))/(cd))

# Chi-square and degrees of freedom
t
df

# Significance test
pchisq(t, df, lower.tail=FALSE)
```

______________________________________________

compare metric to scalar
```{r}

SB0 <- all_models[["M5_scalar.out"]][["summaries"]][["ChiSqM_Value"]]
SB1 <- all_models[["M4_metric.out"]][["summaries"]][["ChiSqM_Value"]]
c0  <- all_models[["M5_scalar.out"]][["summaries"]][["ChiSqM_ScalingCorrection"]]
c1  <- all_models[["M4_metric.out"]][["summaries"]][["ChiSqM_ScalingCorrection"]]
d0  <- all_models[["M5_scalar.out"]][["summaries"]][["ChiSqM_DF"]]
d1  <- all_models[["M4_metric.out"]][["summaries"]][["ChiSqM_DF"]]
df  <- abs(d0-d1)

# Satora-Bentler scaled Difference test equations
cd <- (((d0*c0)-(d1*c1))/(d0-d1))
t  <- (((SB0*c0)-(SB1*c1))/(cd))

# Chi-square and degrees of freedom
t
df

# Significance test
pchisq(t, df, lower.tail=FALSE)
```

______________________________________________

## Invariance short-cut 
```{r, eval=FALSE}

mx  <- mplusObject(
  TITLE = "INVARIANCE SHORT_CUT - DEMO", 
  VARIABLE = 
    "usevar = stolen-rac_fght;
    
     grouping = freelnch (0=freelnch_0 1=freelnch_1); ", 
  
  ANALYSIS = 
  "Estimator = MLR;
   MODEL = CONFIG METRIC SCALAR;",
  
     MODEL = 
    "VICTIM by stolen* t_hurt p_fight hit damaged bullied;
     VICTIM@1; 
     
     NEG_CLIM by unsafe* disrupt gangs rac_fght;
     NEG_CLIM@1;" ,
  
  PLOT = "",
  OUTPUT = "sampstat residual;",
 
  usevariables = colnames(invar_data), 
  rdata = invar_data)

mx_fit <- mplusModeler(mx, 
          dataout=here("invar_short", "invar_short_cut.dat"),
          modelout=here("invar_short", "invar_short_cut.inp"),
          check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

## Invariance Testing (Chi-square values - Chi-Square difference p-values are biased)


                   Number of                   Degrees of
     Model        Parameters      Chi-Square    Freedom     P-Value

     Configural        62            149.315        68       0.0000
     Metric            54            163.312        76       0.0000
     Scalar            46            179.176        84       0.0000

                                               Degrees of
     Models Compared              Chi-Square    Freedom     P-Value

     Metric against Configural        14.759         8       0.0640
     Scalar against Configural        30.022        16       0.0179
     Scalar against Metric            15.444         8       0.0511


______________________________________________

### End

______________________________________________


## References

Hallquist, M. N., & Wiley, J. F. (2018). MplusAutomation: An R Package for Facilitating Large-Scale Latent Variable Analyses in Mplus. Structural equation modeling: a multidisciplinary journal, 25(4), 621-638.

Horst, A. (2020). Course & Workshop Materials. GitHub Repositories, https://https://allisonhorst.github.io/

Lauff, E., & Ingels, S. J. (2014). Education Longitudinal Study of 2002 (ELS: 2002): A First Look at 2002 High School Sophomores 10 Years Later. First Look. NCES 2014-363. National Center for Education Statistics.

Muthén, L.K. and Muthén, B.O. (1998-2017).  Mplus User’s Guide.  Eighth Edition. Los Angeles, CA: Muthén & Muthén

R Core Team (2017). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL http://www.R-project.org/

Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686