---
title: "Higher Order Factors"
author: "Norwegian University of Science and Technology - A Course in `MplusAutomation`"
subtitle: '*Adam Garber*'
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
library(here);library(glue)
```

______________________________________________

## **Outline**

- Prepare data
- Estimate baseline CFA model
- Estimate Higher Order CFA model
- Create fit table for comparison

______________________________________________

DATA SOURCE: This lab exercise utilizes a subset of the HSLS public-use dataset: High School Longitudinal Study of 2009 (Ingels et al., 2011) [$\color{blue}{\text{See website: nces.ed.gov}}$](https://nces.ed.gov/pubsearch/pubsinfo.asp?pubid=2018142)

______________________________________________

## BEGIN: Higher-Order Factors 

______________________________________________

```{r, eval=TRUE}
# load packages
library(MplusAutomation)
library(tidyverse)
library(here)
library(semPlot)
library(gt)
library(DiagrammeR)
```

## Change starting location to folder `11-higher-order`
```{r}

source("rep_functions.R")

change_here(glue("{project_location}/11-higher-order"))

here()

```

______________________________________________

## Prepare data

______________________________________________

## Read in data
```{r eval=TRUE}

data_raw <- read_csv("https://garberadamc.github.io/project-site/data/hsls_fa_data_subset.csv")

```

______________________________________________

## Reverse code for factor interpretation
```{r, eval=TRUE}

hsls_data <- data_raw 

cols = c("S1MPERS1", "S1MPERS2", "S1MUSELI", "S1MUSECL", "S1MUSEJO",
         "S1MTESTS", "S1MTEXTB", "S1MSKILL", "S1MASSEX", "S1MENJNG",
         "S1SPERS1", "S1SPERS2", "S1SUSELI", "S1SUSECL", "S1SUSEJO",
         "S1STESTS", "S1STEXTB", "S1SSKILL", "S1SASSEX", "S1SENJNG")

hsls_data[ ,cols] <-  5 - hsls_data[ ,cols]

```

______________________________________________

## Run a baseline CFA model with 4 factors (for comparison)

______________________________________________

```{r, eval=FALSE}

m.cfa0  <- mplusObject(
  TITLE = "Higher Order FA Models - HSLS SCIENCE", 
  VARIABLE = 
 "usevar = 
 S1SPERS1 S1SPERS2 S1SUSELI S1SUSECL
 S1SUSEJO S1STESTS S1STEXTB S1SSKILL
 S1SASSEX S1SENJNG S1SWASTE S1SBORIN;",
  
  ANALYSIS = 
 "estimator=mlr; ",
  
     MODEL = 
 "SCI_ID BY S1SPERS1* S1SPERS2;
  SCI_ID@1;
  
  SCI_UT BY S1SUSELI* S1SUSECL S1SUSEJO;
  SCI_UT@1;
  
  SCI_EFF BY S1STESTS* S1STEXTB S1SSKILL S1SASSEX ;
  SCI_EFF@1;
  
  SCI_INT BY S1SENJNG* S1SWASTE S1SBORIN;
  SCI_INT@1; ",
  
  PLOT = "type = plot3;",
  OUTPUT = "sampstat standardized residual modindices (3.84);",
 
  usevariables = colnames(hsls_data), 
  rdata = hsls_data)

m.cfa0.fit <- mplusModeler(m.cfa0, 
              dataout=here("2nd_order_FA", "cfa_baseline.dat"),
              modelout=here("2nd_order_FA", "cfa_baseline.inp"),
              check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

## Make a higher-order model path diagram using package {`DiagrammeR`}

______________________________________________

```{r, eval=TRUE}

grViz(" digraph higher_order_path_diagram {

graph [overlap = true, fontsize = 10,   # this is the 'graph' statement
       fontname = Times,
       label= '']    

  node [shape = box]           # this is the 'node' statement
  
  ID1; ID2; UT1; UT2; UT3; UT1; UT2;
  UT3; EFF1; EFF2; EFF3; EFF4; INT1; INT2; INT3;
        
  node [shape = circle, fixedsize = true,
        width = 1.5, label = 'Science Identity'] 
        
  F1;
  
  node [shape = circle, fixedsize = true,
        width = 1.5, label = 'Science Utility'] 
        
  F2;
  
  node [shape = circle, fixedsize = true,
        width = 1.5, label = 'Science Efficacy'] 
        
  F3;
  
  node [shape = circle, fixedsize = true,
        width = 1.5, label = 'Science Interest'] 
        
  F4;
  
  node [shape = circle, fixedsize = true,
        width = 1.5, label = 'Adaptive Science Orientation'] 
        
  G1;
  
  edge [color = black]    # this is the 'edge' statement
  
  F1->{ID1 ID2}
  F2->{UT1 UT2 UT3}
  F3->{EFF1 EFF2 EFF3 EFF4}
  F4->{INT1 INT2 INT3}
  G1->{F1 F2 F3 F4}

}")

```

______________________________________________

## Run a higher-order model model with 4 sub-factors

______________________________________________

```{r, eval=FALSE}

m.cfa1  <- mplusObject(
  TITLE = "Higher Order FA Models - HSLS SCIENCE", 
  VARIABLE = 
 "usevar = 
 S1SPERS1 S1SPERS2 S1SUSELI S1SUSECL
 S1SUSEJO S1STESTS S1STEXTB S1SSKILL
 S1SASSEX S1SENJNG S1SWASTE S1SBORIN;",
  
  ANALYSIS = 
 "estimator=mlr; ",
  
     MODEL = 
 "SCI_ID BY S1SPERS1* S1SPERS2;
  SCI_ID@1;
  
  SCI_UT BY S1SUSELI* S1SUSECL S1SUSEJO;
  SCI_UT@1;
  
  SCI_EFF BY S1STESTS* S1STEXTB S1SSKILL S1SASSEX ;
  SCI_EFF@1;
  
  SCI_INT BY S1SENJNG* S1SWASTE S1SBORIN;
  SCI_INT@1;
 
  ! Regress the higher-order factor on the 4 sub-factors
  F2NDORDR BY SCI_ID SCI_UT SCI_EFF SCI_INT" ,
  
  PLOT = "type = plot3;",
  OUTPUT = "sampstat standardized residual modindices (3.84);",
 
  usevariables = colnames(hsls_data), 
  rdata = hsls_data)

m.cfa1.fit <- mplusModeler(m.cfa1, 
              dataout=here("2nd_order_FA", "cfa_baseline.dat"),
              modelout=here("2nd_order_FA", "cfa_2nd_order.inp"),
              check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

## Generate a higher-order model path diagram from Mplus Output with {`semPlot`}

______________________________________________


```{r}
order2_model <- readModels(here("2nd_order_FA",
                                "cfa_2nd_order.out"))

# plot model:
semPaths(order2_model,
          intercepts=FALSE)
```

______________________________________________

## Compare model fit of baseline and higher-order models

______________________________________________

## Read into R summary of all models
```{r}

models_2 <- readModels(here("2nd_order_FA"), quiet = TRUE)

```

______________________________________________

## Extract relevant data and generate table

```{r}

order2_table <- LatexSummaryTable(models_2, 
  keepCols=c("Filename", "Parameters",
             "ChiSqM_Value", "CFI","TLI",
             "SRMR", "RMSEA_Estimate",
             "RMSEA_90CI_LB", "RMSEA_90CI_UB"), 
  sortBy = "Filename")

order2_table %>%
  mutate(Filename = str_remove(Filename, ".out")) %>% 
  gt() %>% 
  cols_label(
    Filename = "Model",
    Parameters = "Par",
    ChiSqM_Value = "ChiSq",
    CFI = "CFI", TLI = "TLI", SRMR = "SRMR",
    RMSEA_Estimate = "RMSEA",
    RMSEA_90CI_LB = "Lower CI",
    RMSEA_90CI_UB = "Upper CI")

```

______________________________________________

## References

Hallquist, M. N., & Wiley, J. F. (2018). MplusAutomation: An R Package for Facilitating Large-Scale Latent Variable Analyses in Mplus. Structural equation modeling: a multidisciplinary journal, 25(4), 621-638.

Horst, A. (2020). Course & Workshop Materials. GitHub Repositories, https://https://allisonhorst.github.io/

Ingels, S. J., Pratt, D. J., Herget, D. R., Burns, L. J., Dever, J. A., Ottem, R., ... & Leinwand, S. (2011). High School Longitudinal Study of 2009 (HSLS: 09): Base-Year Data File Documentation. NCES 2011-328. National Center for Education Statistics.

Muthén, L.K. and Muthén, B.O. (1998-2017).  Mplus User’s Guide.  Eighth Edition. Los Angeles, CA: Muthén & Muthén

R Core Team (2017). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL http://www.R-project.org/

Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686