---
title: "Path Models | Indirect Effects | Single Indicator Factors"
author: "Norwegian University of Science and Technology - A Course in `MplusAutomation`"
subtitle: '*Adam Garber*'
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
library(here);library(glue);library(linguisticsdown);library(tidyverse);library(gt)
```

______________________________________________

```{r, echo=FALSE, eval=TRUE, out.width = "100%", out.height= "100%", fig.pos="h"}
include_graphics2("https://garberadamc.github.io/project-site/figures/f1_path_models.png")
```

______________________________________________

## DATA SOURCE: This lab exercise utilizes the NCES public-use dataset: Education Longitudinal Study of 2002 (Lauff & Ingels, 2014) [$\color{blue}{\text{See website: nces.ed.gov}}$](https://nces.ed.gov/surveys/els2002/avail_data.asp)

______________________________________________

## Tools for reproducibility:

```{r, eval=TRUE, echo=FALSE}

lab_tools <- tribble(
  ~"Tool/Package", ~"Purpose/Utility",  ~"Advantages",
 #---------------|-------------------|----------------|,
 "{MplusAutomation} " ,  "Current capabilities supporting full SEM modeling" , "High flexibility",
 "RProject" ,  "Unbreakable file paths & neatness" , "Reproducibility (kindness to your future self)",
 "{tidyverse}" ,  "Intuitive/descriptive function names" , "Accessibility to new users",
 "{here}" , "Unbreakable/consistent file paths across OS" , "Reproducibility (for Science's sake!)",
 "{haven}" ,  "View-able metadata in R from SPSS data-files" , "Getting to know your measures",
 "{ggplot2}" ,  "Clear, customizable, reproducible figures" , "Publication quality data visualizations",
 "pipe operator (%>%)", "Ease of reading/writing scripts", "e.g., first() %>% and_then() %>% and_finally()"
)

lab_tools %>% gt()
```

______________________________________________
## Load packages
```{r, eval=TRUE}

library(MplusAutomation)
library(haven)
library(tidyverse)
library(here)
library(corrplot)
library(janitor)
library(ggridges)
library(DiagrammeR)
library(semPlot)
library(sjPlot)
library(gt)
```

______________________________________________

## Change starting location to folder `12-path-models`
```{r}

source("rep_functions.R")

change_here(glue("{project_location}/12-path-models"))

here()

```

______________________________________________

## Read in SPSS data
```{r, eval=TRUE}

spss_data <- read_spss("https://garberadamc.github.io/project-site/data/els_sub1_spss.sav") %>% 
  janitor::clean_names() # makes all variable names "lowercase snakecase"

```

## Preparations: subset, rename, and reorder columns

1. **subset:** select columns in 3 ways, remove columns with (-), select by index number, and select by column name 
2. **rename:** change variable names to be descriptive and within the Mplus 8 character limit
3. **reorder:** this makes it easy to choose sequential variables for {MplusAutomation}

## Remove colums with `select(-column_removed)`
```{r}
spss_sub0 <- spss_data %>% 
  select(-stu_id,   -sch_id,   -byrace,     
         -byparace, -byparlng, -byfcomp,   
         -bypared,  -bymothed, -byfathed,
         -bysctrl,  -byurban,  -byregion)
```

## Select the first 9 columns (by index) and select the next 17 columns (by name)
```{r}
spss_sub1 <- spss_sub0 %>% 
  select(1:9,
         bys20a, bys20h, bys20j, bys20k, bys20m, bys20n,
         bys21b, bys21d, bys22a, bys22b, bys22c, bys22d,
         bys22e, bys22g, bys22h, bys24a, bys24b) %>% 
  rename("stu_exp" = "bystexp",   # "NEW_NAME" = "OLD_NAME"
         "par_asp" = "byparasp",
         "mth_read" = "bytxcstd",
         "mth_test" = "bytxmstd",
         "rd_test" = "bytxrstd",
         "freelnch" = "by10flp",
         "stu_tch" = "bys20a",
         "putdownt" = "bys20h",
         "unsafe" = "bys20j",
         "disrupt" = "bys20k",
         "gangs" = "bys20m",
         "rac_fght" = "bys20n",
         "fair" = "bys21b",
         "strict" = "bys21d",
         "stolen" = "bys22a",
         "drugs" = "bys22b",
         "t_hurt" = "bys22c",
         "p_fight" = "bys22d",
         "hit" = "bys22e",
         "damaged" = "bys22g",
         "bullied" = "bys22h",
         "late" = "bys24a",
         "skipped" = "bys24b")
```

______________________________________________

## More housekeeping: reorder columns 
```{r}

spss_sub2 <-  spss_sub1 %>% 
  select(
    bystlang,                                        # dichotomous (yes,no)
    freelnch, byincome,                              # ordinal (binned, continuous scale)
    stolen, t_hurt, p_fight, hit, damaged, bullied,  # ordinal frequency (3-point)
    unsafe, disrupt, gangs, rac_fght,                # ordinal Likert (4-point scale)
    late, skipped,                                   # ordinal frequency (4-point scale)
    mth_test, rd_test)                               # continuous (standardized test scores)


```

______________________________________________

## Make a codebook including metadata using {`sjPlot`}

```{r, eval=FALSE}

sjPlot::view_df(spss_sub2)

```

______________________________________________

## Types of data for different tasks

- `SAV` (e.g., spss_data.sav): this data format is for SPSS files & contains variable labels (meta-data)
- `CSV` (e.g., r_ready_data.csv): this is the preferable data format for reading into R (no labels)
- `DAT` (e.g., mplus_data.dat): this is the data format used to read into Mplus (no column names or strings)

NOTE: Mplus also accepts `.txt` formatted data (e.g., `mplus_data.txt`)

______________________________________________

## Converting data between 3 formats: writing and reading data

______________________________________________

Write a CSV datafile (preferable format for reading into R, with SPSS labels removed)
```{r, eval=FALSE}

write_csv(spss_sub2, here("data", "els_sub6_data.csv"))

```

Read the unlabeled data back into R
```{r,eval=TRUE}

tidy_data <- read_csv(here("data", "els_sub6_data.csv"))

```

Write a DAT datafile for Mplus (this function removes header row & converts missing values to non-string)
```{r, eval=FALSE}

prepareMplusData(tidy_data, here("data", "els_sub6_data.dat"))

```

______________________________________________

## Make a `tribble` table

```{r, eval=TRUE}

var_table <- tribble(
   ~"Name",    ~"Labels",                                      ~"Value Labels (limit)",  
 #-----------|-----------------------------------------------|--------------------------------------|,
  "bystlang" , "Whether English is students native language" ,"0=No, 1=Yes", 
  "freelnch" , "Grade 10 percent free lunch-categorical"     ,"0=0-5%, 7=76-100%", 
  "byincome" , "Total family income from all sources 2001"   ,"1=None, 13=$200,001 or more", 
  "stolen"   , "Had something stolen at school"              ,"1=Never, 3=More than twice", 
  "t_hurt"   , "Someone threatened to hurt 10th grader at school","1=Never, 3=More than twice", 
  "p_fight"  , "Got into a physical fight at school"         ,"1=Never, 3=More than twice"  ,
  "hit"      , "Someone hit 10th grader"                     ,"1=Never, 3=More than twice"  , 
  "damaged"  , "Someone damaged belongings"                  ,"1=Never, 3=More than twice"  , 
  "bullied"  , "Someone bullied or picked on 10th grader"    ,"1=Never, 3=More than twice"  , 
  "unsafe"   , "Does not feel safe at this school"           ,"1=Strongly agree, 4=Strongly disagree"  , 
  "disrupt"  , "Disruptions get in way of learning"          ,"1=Strongly agree, 4=Strongly disagree"  , 
  "gangs"    , "There are gangs in school"                   ,"1=Strongly agree, 4=Strongly disagree"  , 
  "rac_fght" , "Racial-ethnic groups often fight"            ,"1=Strongly agree, 4=Strongly disagree"  ,
  "late"     , "How many times late for school"              ,"1=Never, 4=10 or more times"  , 
  "skipped"  , "How many times cut-skip classes"             ,"1=Never, 4=10 or more times"  ,
  "mth_test" , "Math test standardized score"                ,"0-100"  , 
  "rd_test"  , "Reading test standardized score"             ,"0-100"  , 
)

var_table %>% gt()
```


______________________________________________

## Take a look at the data - some practice with {`ggplot2`}

______________________________________________

Make a faceted box plot
```{r}

# some formatting, add labels to `bystlang` for plot 
tidy_data <- tidy_data %>% 
  mutate(
    bystlang = factor(bystlang, 
    labels = c(`0` = "Non-English", `1` = "English")))


  ggplot(data=drop_na(tidy_data), aes(y=mth_test)) +
    geom_boxplot() +
    facet_wrap(~bystlang) +
    labs(x = "Native language",
         y = "Math test (standardized score)")
    
    
```

Make a density plot
```{r}

  ggplot(data=drop_na(tidy_data), aes(x=mth_test)) +
    geom_density(aes(fill = bystlang),
                 color = NA,
                 show.legend = FALSE) +
    facet_wrap(~bystlang) +
    theme_light()
    
```

Ridgeline plot {`ggridges`}  
```{r}

# A ridgeline plot is good way to compare distributions across groups.
# In the plot below the distribution of reading test scores is grouped
# by level of the freelunch variable. 

ridge_graph <- ggplot(data = drop_na(tidy_data),
                      aes(x = rd_test, y = factor(freelnch))) +
  geom_density_ridges(aes(fill = factor(freelnch)),
                      size = 0.2,
                      alpha = 0.7,
                      show.legend = FALSE) +
  scale_x_continuous(lim = c(0,100)) +
    scale_y_discrete(lim = levels(tidy_data$freelnch), 
                   labels = c("0-5%", "6-10%", "11-20%","21-30%",
                              "31-50%", "51-75%", "76-100%")) +
    labs(x = "Reading test (standardized score)",
         y = "Percent Free Lunch",
     title = "Grade 10 Reading Test Scores by Percent Free Lunch in School",
  subtitle = "Source: ElS 2002") +
  theme_minimal() 

ridge_graph

```

______________________________________________

## Look at all bivariate relations

______________________________________________

```{r}

t_cor <- cor(tidy_data[,4:17], use = "pairwise.complete.obs")

corrplot(t_cor, 
         method = "color",
         type = "upper", 
         tl.col="black", 
         tl.srt=45)
```

______________________________________________

## Run some path models with {MplusAutomation}

______________________________________________

Practice run, use `type=basic` to get descriptives
```{r, eval=FALSE}

m_basic  <- mplusObject(
  TITLE = "RUN TYPE = BASIC ANALYSIS - LAB 1", 
  VARIABLE = 
    " ! an mplusObject() will always need a 'usevar' statement 
      ! ONLY specify variables that will be used in analysis
      ! lines of code in MPLUS ALWAYS end with a semicolon ';'
      
    usevar =
    bystlang freelnch byincome stolen t_hurt p_fight
    hit damaged bullie, unsafe disrupt gangs rac_fght
    late skipped mth_test rd_test;",          
  
  ANALYSIS = 
    "type = basic" ,
  
  MODEL = "" ,
  
  PLOT = "",
  
  OUTPUT = "",
  
  usevariables = colnames(tidy_data),   # tell MplusAutomation the column names to use
  rdata = tidy_data)                    # this is the data object used (must be un-label)

m_basic_fit <- mplusModeler(m_basic,
               dataout=here("mplus_files", "nces.dat"),
               modelout=here("mplus_files", "m0_basic_nces.inp"),
               check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

## Run a path model with model indirect (to estimate the indirect effect) 

______________________________________________

```{r, echo=FALSE, eval=TRUE, out.width = "65%", out.height= "65%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/model_indirect.png")

```

*Figure 1*. Path Diagram of Multiple Indirect Paths Model

______________________________________________

Visualize the path diagram using the {`DiagrammeR`} package
```{r,eval=FALSE}

mermaid("
graph LR
  bystlang-->late
  bystlang-->skipped
  bystlang-->mth_test
  late-->skipped
  late-->mth_test
  skipped-->mth_test
")
```

______________________________________________

Run model depicted above with multiple indirect paths
```{r, eval=FALSE}

m1_ind  <- mplusObject(
  TITLE = "m1 model indirect - Lab 1", 
  VARIABLE = 
   "usevar =
    bystlang       ! covariate
    late skipped   ! mediators
    mth_test;      ! outcome   ",          
  
  ANALYSIS = 
    "estimator = MLR" ,
  
  MODEL = 
   "late on bystlang ;
    skipped on late bystlang ;
    mth_test on late skipped bystlang;
    
    Model indirect:
    mth_test ind bystlang;
    mth_test via late skipped bystlang; " ,
  
  OUTPUT = "sampstat standardized",
  
  usevariables = colnames(tidy_data),   
  rdata = tidy_data)                    

m1_ind_fit <- mplusModeler(m1_ind,
              dataout=here("mplus_files", "nces.dat"),       
              modelout=here("mplus_files", "m1_indirect_nces.inp"),
              check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

Generate a path diagram from Mplus output with {`semPlot`}

```{r}
order2_model <- readModels(here("mplus_files",
                                "m1_indirect_nces.out"))

# plot model:
semPaths(order2_model,
          intercepts=FALSE)
```

______________________________________________

## Single indicator factors

______________________________________________

Model specifications:

- Fix the loading to 1
- Then fix the residual variance to a specific value (you are not estimating a measurement parameter)

Using reliability you fix the residual variance at:

$(1‐reliabilty) *variance$

______________________________________________

Lab example of single indicator factor model:

```{r, echo=FALSE, eval=TRUE, out.width = "65%", out.height= "65%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/single_indicator_factor.png")

```

*Figure 2*. Path Diagram of Single Indicator Factor Model

______________________________________________

create a mean score variable called `mean_score`
```{r}

tidy_data2 <- tidy_data %>%
  mutate(mean_scr = rowSums(select(., late:skipped))/2)

```

    - Reliability = .8 (set to)
    - Variance = .77 (mean_score)

Function to fix the residual variance 
```{r}

resid_var <- function(reliability, variance) {  
  
 y <- ((1-reliability)*variance)
    return(y)
}

(y01 <- resid_var(reliability = .8, variance = .77))
```

______________________________________________

Run model with single indicator factor
```{r, eval=FALSE}

m2_sif  <- mplusObject(
  TITLE = "m2 single indicator factor - Lab 1", 
  VARIABLE = 
   "usevar =
    unsafe disrupt gangs rac_fght  ! factor 1
    mth_test                       ! outcome
    mean_scr;                      ! mediator ",                  
  
  ANALYSIS = 
    "estimator = MLR" ,
  
  MODEL = 
   "! measurement model
    factor1 by unsafe, disrupt, gangs, rac_fght; 
    
    SIF by mean_scr@1; ! fix factor loading to 1

    mean_scr@.154;     ! fix residual variance
    
    ! structural model
    mth_test on factor1 SIF;
    SIF on factor1; ",
  
  OUTPUT = "sampstat standardized",
  
  usevariables = colnames(tidy_data2),   
  rdata = tidy_data2)                    

m2_sif_fit <- mplusModeler(m2_sif,
                    dataout=here("mplus_files", "nces.dat"),       
                    modelout=here("mplus_files", "m2_sif_nces.inp"),
                    check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

Generate a path diagram from Mplus output with {`semPlot`}

```{r}
order2_model <- readModels(here("mplus_files",
                                "m2_sif_nces.out"))

# plot model:
semPaths(order2_model,
          intercepts=FALSE)
```

```{r, echo=FALSE, eval=TRUE, out.width = "85%", out.height= "85%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/f2_single_indicator.png")

```


______________________________________________


## References

Hallquist, M. N., & Wiley, J. F. (2018). MplusAutomation: An R Package for Facilitating Large-Scale Latent Variable Analyses in Mplus. Structural equation modeling: a multidisciplinary journal, 25(4), 621-638.

Horst, A. (2020). Course & Workshop Materials. GitHub Repositories, https://https://allisonhorst.github.io/

Muthén, L.K. and Muthén, B.O. (1998-2017).  Mplus User’s Guide.  Eighth Edition. Los Angeles, CA: Muthén & Muthén

R Core Team (2017). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL http://www.R-project.org/

Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686
