---
title: "Competing Path Models"
author: "Norwegian University of Science and Technology - A Course in `MplusAutomation`"
subtitle: '*Adam Garber*'
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
library(here);library(glue);library(linguisticsdown)
```

______________________________________________

## Preparation

______________________________________________


Example of competing path models study from [$\color{blue}{\text{Nishina, Juvonen, Witkow (2005)}}$](https://www.tandfonline.com/doi/pdf/10.1207/s15374424jccp3401_4)

```{r, echo=FALSE, eval=TRUE, out.width = "100%", out.height= "100%", fig.pos="h"}
include_graphics2("https://garberadamc.github.io/project-site/figures/fig1_Nishina.png")
```
*figure*. Picture adapted from Nishina, Juvonen, Witkow (2005)

______________________________________________

## Data source: 

This lab exercise utilizes the *California Test Score Data Set 1998-1999* from the California Department of Education (Stock, James, and Watson, 2003) [$\color{blue}{\text{See documentation here}}$](https://vincentarelbundock.github.io/Rdatasets/doc/Ecdat/Caschool.html)

This dataset is available via the R-package {`Ecdat`} and can be directly loaded into the R environment. 

**Note:** All models specified in the following exercise are for demonstration only and are **not** theoretically justified or valid. 

______________________________________________

Load packages
```{r, eval=TRUE}

library(MplusAutomation)
library(haven)
library(tidyverse)
library(here)
library(corrplot)
library(reshape2)
library(janitor)
library(ggridges)
library(DiagrammeR)
library(semPlot)
library(sjPlot)
library(Ecdat)
library(gt)
library(gtsummary)
```

______________________________________________

## Change starting location to folder `13-comp-models`
```{r}

source("rep_functions.R")

change_here(glue("{project_location}/13-comp-models"))

here()

```

______________________________________________

## Begin lab 2 exercise

______________________________________________

Read the dataframe into your R-environment from package {`Ecdat`}
```{r}

data(Caschool)

ca_schools <- as.data.frame(Caschool)

```

Look at the data with `glimpse`

```{r, eval=FALSE}

glimpse(ca_schools)

```

______________________________________________

Subset variables to use in path model analyses with `select`
```{r}

path_vars <- ca_schools %>% 
  select(str, expnstu, compstu, elpct, mealpct,
         readscr, mathscr, testscr)

```

______________________________________________

## Explore the data

K through 8th grade schools in California ($N = 420$)

Take a look at focal variables, make a `tribble` table
```{r, eval=TRUE}

var_table <- tribble(
   ~"Name",    ~"Labels",                                     
 #-----------|----------------------------------------------|,
  "str"       , "student teacher ratio"                      ,
  "expnstu"   , "expenditure per student"                    ,
  "compstu"   , "computer per student"                       ,
  "elpct"     , "percent of English learners"                ,
 "mealpct"    , "percent qualifying for reduced-price lunch" ,
  "readscr"   , "average reading score"                      ,
  "mathscr"   , "average math score"                         ,
  "testscr"   , "average test score (read.scr+math.scr)/2"   )

var_table %>% 
  gt()
```

______________________________________________

check some basic descriptives with the {`gtsummary`} package
```{r}

table1 <- tbl_summary(path_vars,
                      statistic = list(all_continuous() ~ "{mean} ({sd})"),
                      missing = "no" ) %>%
  bold_labels() 

table1
```

______________________________________________

look at shape of variable distributions
```{r}

melt(path_vars) %>%                  
  ggplot(., aes(x=value, label=variable)) +   
  geom_density(aes(fill = variable),
               alpha = .5, show.legend = FALSE) + 
  facet_wrap(~variable, scales = "free")  +
  theme_minimal()

```

______________________________________________

look at correlation matrix with {`corrplot`}

```{r}

p_cor <- cor(path_vars, use = "pairwise.complete.obs")

corrplot(p_cor, 
         method = "color",
         type = "upper", 
         tl.col="black", 
         tl.srt=45)

```

______________________________________________

## Specifying path models using {`MplusAutomation`}

recall what the unrestricted variance-covariance matrix **looks** like 

```{r, echo=FALSE, eval=TRUE, out.width = "80%", out.height= "80%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/unrestrict_model.png")
```

*figure*. Unrestricted variance covariance matrix picture from {`openMX`} video tutorial. 

______________________________________________

## Estimate model 1 

Indirect path model:

1. **covariate**: ratio of computers to students (`compstu`)
2. **mediator**: percent qualifying for reduced-price lunch (`mealpct`)
3. **outcome**: average math score (`mathscr`)

______________________________________________

Path diagram model 1
```{r, echo=FALSE, eval=TRUE, out.width = "80%", out.height= "80%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/model1_indirect.png")
```

______________________________________________

```{r, eval=FALSE}

m1_path  <- mplusObject(
  TITLE = "m1 model indirect - Lab 1", 
  VARIABLE = 
   "usevar =
    compstu         ! covariate
    mealpct         ! mediator 
    mathscr;        ! outcome",            
  
  ANALYSIS = 
    "estimator = MLR" ,
  
  MODEL = 
   "mathscr on compstu;         ! direct path (c')
    mathscr on mealpct;         ! b path
    mealpct on compstu;         ! a path
    
    Model indirect:
    mathscr ind compstu;" ,
  
  OUTPUT = "sampstat standardized modindices (ALL)",
  
  usevariables = colnames(path_vars),   
  rdata = path_vars)                    

m1_path_fit <- mplusModeler(m1_path,
               dataout=here("mplus_files", "CA_schls.dat"),       
               modelout=here("mplus_files", "m1_path_indirect.inp"),
               check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

View path diagram for model 1 with standardized estimates (using Diagrammer in Mplus)
```{r, echo=FALSE, eval=TRUE, out.width = "80%", out.height= "80%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/m1_path.png")

```

______________________________________________

## Estimate model 2

change variable status (**switch mediator and covariate variables**)

Indirect path model:

1. covariate: percent qualifying for reduced-price lunch (`mealpct`)
2.  mediator: ratio of computers to students (`compstu`)
3.   outcome: average math score (`mathscr`)

______________________________________________

Path diagram model 2
```{r, echo=FALSE, eval=TRUE, out.width = "80%", out.height= "80%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/model2_indirect.png")
```

______________________________________________


```{r, eval=FALSE}

m2_path  <- mplusObject(
  TITLE = "m2 model indirect", 
  VARIABLE = 
   "usevar =
    mealpct           ! covariate
    compstu           ! mediator 
    mathscr;          ! outcome",            
  
  ANALYSIS = 
    "estimator = MLR" ,
  
  MODEL = 
   "mathscr on compstu;         ! direct path (c')
    mathscr on mealpct;         ! b path
    mealpct on compstu;         ! a path
    
    Model indirect:
    mathscr ind compstu;" ,
  
  OUTPUT = "sampstat standardized modindices (ALL)",
  
  usevariables = colnames(path_vars),   
  rdata = path_vars)                    

m2_path_fit <- mplusModeler(m2_path,
                     dataout=here("mplus_files", "CA_schls.dat"),       
                    modelout=here("mplus_files", "m2_path_indirect.inp"),
                    check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

View path diagram for model 2 with standardized estimates (using the Diagrammer in Mplus)
```{r, echo=FALSE, eval=TRUE, out.width = "80%", out.height= "80%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/m2_path.png")

```

______________________________________________

## Estimate model 3 

Path model with interaction (moderation):

1. covariate-moderator: percent qualifying for reduced-price lunch (`mealpct`)
2. covariate-moderator: ratio of computers to students (`compstu`)
3. outcome: average math score (`mathscr`)

______________________________________________

Path diagram model 3
```{r, echo=FALSE, eval=TRUE, out.width = "80%", out.height= "80%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/model3_interact.png")
```

______________________________________________

```{r, eval=FALSE}

m3_path  <- mplusObject(
  TITLE = "m3 model interact", 
  VARIABLE = 
   "usevar =
    compstu           ! covariate-moderator
    mealpct           ! covariate-moderator
    mathscr           ! outcome
    int_ab;           ! interaction term ", 
  
  DEFINE = 
    "int_ab = compstu*mealpct;  ! create interaction term" ,
  
  ANALYSIS = 
    "estimator = MLR" ,
  
  MODEL = 
   "mathscr on compstu mealpct int_ab; ",
  
  OUTPUT = "sampstat standardized modindices (ALL)",
  
  usevariables = colnames(path_vars),   
  rdata = path_vars)                    

m3_path_fit <- mplusModeler(m3_path,
                    dataout=here("mplus_files", "CA_schls.dat"),       
                    modelout=here("mplus_files", "m3_path_interact.inp"),
                    check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

View path diagram for model 3 with standardized estimates (using the Diagrammer in Mplus)
```{r, echo=FALSE, eval=TRUE, out.width = "80%", out.height= "80%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/m3_path.png")

```

______________________________________________

## Estimate model 4

______________________________________________

Two indirect paths or "mediators"
```{r, eval=FALSE}

m4_path  <- mplusObject(
  TITLE = "m4 model indirect (two mediators)", 
  VARIABLE = 
   "usevar =
    str               ! covariate
    elpct             ! mediator
    mealpct           ! mediator
    mathscr           ! outcome", 
  
  DEFINE = 
    "int_ab = compstu*mealpct;  ! create interaction term" ,
  
  ANALYSIS = 
    "estimator = MLR" ,
  
  MODEL = 
   "mathscr on str;             ! direct path (c')
    mathscr on elpct mealpct;   ! b paths
    elpct mealpct on str;       ! a paths
    
    Model indirect:
    mathscr ind str;" ,
  
  OUTPUT = "sampstat standardized modindices (ALL)",
  
  usevariables = colnames(path_vars),   
  rdata = path_vars)                    

m4_path_fit <- mplusModeler(m4_path,
                    dataout=here("mplus_files", "CA_schls.dat"),       
                    modelout=here("mplus_files", "m4_path_indirect.inp"),
                    check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

View path diagram for model 4 with standardized estimates (using the Diagrammer in Mplus)
```{r, echo=FALSE, eval=TRUE, out.width = "80%", out.height= "80%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/m4_path.png")

```

______________________________________________

## Estimate model 5

______________________________________________

Add modification statement - correlate mediators `mealpct with elpct`
```{r, eval=FALSE}

m5_path  <- mplusObject(
  TITLE = "m5 model indirect (correlate mediators)", 
  VARIABLE = 
   "usevar =
    str               ! covariate
    elpct             ! mediator
    mealpct           ! mediator
    mathscr           ! outcome", 
  
  DEFINE = 
    "int_ab = compstu*mealpct;  ! create interaction term" ,
  
  ANALYSIS = 
    "estimator = MLR" ,
  
  MODEL = 
   "mathscr on str;             ! direct path (c')
    mathscr on elpct mealpct;   ! b paths
    elpct mealpct on str;       ! a paths
    
    mealpct with elpct          ! modification statement 
    
    Model indirect:
    mathscr ind str; " ,
  
  OUTPUT = "sampstat standardized modindices (ALL)",
  
  usevariables = colnames(path_vars),   
  rdata = path_vars)                    

m5_path_fit <- mplusModeler(m5_path,
                    dataout=here("mplus_files", "CA_schls.dat"),       
                    modelout=here("mplus_files", "m5_path_indirect.inp"),
                    check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

View path diagram for model 5 with standardized estimates (using the Diagrammer in Mplus)
```{r, echo=FALSE, eval=TRUE, out.width = "80%", out.height= "80%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/m5_path.png")

```



______________________________________________

## End 

______________________________________________

## References

Hallquist, M. N., & Wiley, J. F. (2018). MplusAutomation: An R Package for Facilitating Large-Scale Latent Variable Analyses in Mplus. Structural equation modeling: a multidisciplinary journal, 25(4), 621-638.

Horst, A. (2020). Course & Workshop Materials. GitHub Repositories, https://https://allisonhorst.github.io/

Ingels, S. J., Pratt, D. J., Herget, D. R., Burns, L. J., Dever, J. A., Ottem, R., ... & Leinwand, S. (2011). High School Longitudinal Study of 2009 (HSLS: 09): Base-Year Data File Documentation. NCES 2011-328. National Center for Education Statistics.

Muthén, L.K. and Muthén, B.O. (1998-2017).  Mplus User’s Guide.  Eighth Edition. Los Angeles, CA: Muthén & Muthén

R Core Team (2017). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL http://www.R-project.org/

Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686





