---
title: "Moderation"
author: "Norwegian University of Science and Technology - A Course in `MplusAutomation`"
subtitle: '*Adam Garber*'
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
library(here);library(glue);library(linguisticsdown)
```

______________________________________________

## Lab preparation

______________________________________________


## Data source for example 1: 

The first example utilizes the *Vocabulary and Education* dataset from the National Opinion Research Center General Social Survey. GSS Cumulative Datafile 1972-2016 (Fox, 2008) [$\color{blue}{\text{See documentation here}}$](https://vincentarelbundock.github.io/Rdatasets/doc/carData/Vocab.html)

This dataset is available via the R-package {`carData`} and can be directly loaded into the R environment.

**Note:** All models specified in the following exercise are for demonstration only and are **not** theoretically justified or valid. 

______________________________________________

```{r, eval=FALSE}
# equatiomatic is not yet on CRAN. Install the development version from GitHub with
remotes::install_github("datalorax/equatiomatic", force = TRUE)
```

```{r, eval=TRUE}

library(tidyverse)
library(MplusAutomation)
library(rhdf5)
library(here)
library(gt)
library(gtsummary)
library(carData)
library(Ecdat)

```

______________________________________________

## Change starting location to folder `14-moderation`
```{r}

source("rep_functions.R")

change_here(glue("{project_location}/14-moderation"))

here()

```

______________________________________________

## Begin lab 2 exercise

______________________________________________

Read the dataframe into your R-environment from package {`carData`}
```{r}

data(Vocab)

vocab <- as.data.frame(Vocab)

```

______________________________________________

Take a look at focal variables, make a `tribble` table
```{r, eval=TRUE, echo=FALSE, results='asis'}

var_table <- tribble(
   ~"Name",      ~"Labels",                                     
 #--------------|--------------------------------|,
  "year"        , "Year of the survey"         ,
  "sex"         , "Sex of the respondent (Female or Male)"       ,
  "education"   , "Education, in years"          ,
  "vocabulary"  , "Vocabulary test score: number correct on a 10-word test"   )

var_table %>% gt()
```

______________________________________________

check some basic descriptives with the {`gtsummary`} package
```{r}

table1 <- tbl_summary(vocab,
                      statistic = list(all_continuous() ~ "{mean} ({sd})"),
                      missing = "no" ) %>%
  bold_labels() 

table1
```

______________________________________________

## Estimate moderation example 1 

1. covariate: Years of education (`education`)
2. moderator: Year of the survey with 2-levels 1974 and 2016 (`year`)
3. outcome: Vocabulary test score - number correct on a 10-word test (`vocabulary`)

______________________________________________

```{r, echo=FALSE, eval=TRUE, out.width = "80%", out.height= "80%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/lab3_model1.png")
```

______________________________________________

```{r, eval=FALSE}
m1_model  <- mplusObject(
  TITLE = "m1 model moderation", 
  VARIABLE = 
   "usevar =
    year              ! covariate/moderator
    education         ! covariate
    vocabulary        ! outcome
    int_yred;         ! interaction of year and education", 
  
  DEFINE = 
    "center education (grandmean);
     int_yred = year*education;  ! create interaction term ",
  
  ANALYSIS = 
    "estimator = MLR" ,
  
  MODEL = 
   "[vocabulary](b0);
    vocabulary on
    year(b1)
    education(b2)
    int_yred(b3); " ,
  
  MODELCONSTRAINT =  
  "LOOP(x,-1,1,0.01);
   PLOT(y1974 y2016);
   new(hi_y1974 lo_y1974 hi_y2016 lo_y2016 diff_hi); 
   y1974 = b0 + b2*x;
   y2016 = b0 + b1 + (b2+b3)*x;
   
   hi_y1974 = b0 + b2*(6.28);  !!! education 2SD = 2*sqrt(9.85) = 6.28 !!!
   lo_y1974 = b0 + b2*(-6.28);
   hi_y2016 = b0 + b1 + (b2 + b3)*(6.28);
   lo_y2016 = b0 + b1 + (b2 + b3)*(-6.28);
    diff_hi = hi_y2016 - hi_y1974; ",
  
  OUTPUT = "sampstat standardized modindices (3.84)",
  
  PLOT = "type=plot3;",
  
  usevariables = colnames(vocab2),   
  rdata = vocab2)                    

m1_model_fit <- mplusModeler(m1_model,
                    dataout=here("mplus_files", "vocab.dat"),       
                    modelout=here("mplus_files", "model1_vocab.inp"),
                    check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

## Create the simple slope plot from Mplus model output 

______________________________________________

Extract the output parameters generated using the `model constraint`
```{r, eval=FALSE}

 simp_slope <- data.frame(m1_model_fit[["results"]][["parameters"]][["unstandardized"]]) %>% 
   filter(paramHeader == "New.Additional.Parameters")  %>%
   filter(param!= "DIFF_HI") %>% 
   select(param, est, se) %>% 
   mutate(year = case_when(
      param %in% c("HI_Y1974", "LO_Y1974") ~ "1974",
      param %in% c("HI_Y2016", "LO_Y2016") ~ "2016")) %>% 
   mutate(education = case_when(
      param %in% c("HI_Y1974", "HI_Y2016") ~ 6.28,
      param %in% c("LO_Y1974", "LO_Y2016") ~ -6.28)) 
   
```

Plot the interaction effect with `ggplot` using theme from {`hrbrthemes`} package
```{r, eval=FALSE}
# un-center 'education' so values on x-axis are on the original scale
plot_data <- simp_slope %>% mutate(education = education + 12.9)

ggplot(plot_data, aes(x=education, y=est, color=year, group=year)) +
  geom_point(size=0) +
  geom_line() +
  geom_errorbar(aes(ymin=est-se, ymax=est+se),
                width=.2) +
  scale_x_continuous(breaks = c(seq(6,20,2))) +
  labs(title = "Simple Slopes Graph",
       subtitle = "Vocabulary test score predicted by years of education in 1974 & 2016",
           x = "Education (years)",
           y = "Vocabulary test score") +
  theme_minimal()
```


```{r, eval=FALSE}
ggsave(here("figures", "m1_simple_slope.png"), height = 6, width = 8)
```

```{r, echo=FALSE, eval=TRUE, out.width = "80%", out.height= "80%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/m1_simple_slope.png")
```

______________________________________________

## Data source for example 2: 

The next example utilizes the **Effects on Learning of Small Class Sizes (Star)** dataset from the *Introduction to Econometrics* textbook. (Stock et al., 2003) [$\color{blue}{\text{See documentation here}}$](https://vincentarelbundock.github.io/Rdatasets/doc/Ecdat/Star.html)

This dataset is available via the R-package {`Ecdat`} and can be directly loaded into the R environment.

______________________________________________

Read the dataframe into your R-environment from package {`Ecdat`}
```{r}

data(Star)

star_data <- as.data.frame(Star)

```


Take a look at the variables in the `Star` dataset
```{r, eval=TRUE, echo=FALSE, results='asis'}

var_table <- tribble(
   ~"Name",      ~"Labels",                                    
 #--------------|--------------------------------|,
  "tmathssk"    , "total math scaled score"       ,
  "treadssk"    , "total reading scaled score"    ,
  "classk"      , "type of class (small, regular, regular with aide)",
  "totexpk"     , "years of total teaching experience")

var_table %>% gt()
```

Subset and recode variables to use in moderation model with `select`, `mutate`, and `case_when`
```{r}

mod_data <- star_data %>% 
  select(totexpk, # years of total teaching experience
         classk,  # type of class, a factor with levels (regular,small.class,regular.with.aide)
         tmathssk, treadssk)  %>% 
  mutate(classk = case_when(
         classk == "small.class" ~ "small.class",
         classk %in% c("regular.with.aide", "regular") ~ "regular")) %>% 
  mutate(classk = fct_rev(classk))

```

```{r, echo=FALSE, eval=TRUE, out.width = "80%", out.height= "80%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/Lab3_model2.png")
```

______________________________________________

## Estimate moderation example 2 

1. covariate: Years of education (`totexpk`)
2. moderator: type of class (small, regular) (`classk`)
3. outcome 1: total math scaled score (`tmathssk`)
4. outcome 2: total reading scaled score (`treadssk`)

______________________________________________


```{r, eval=FALSE}
m2_model  <- mplusObject(
  TITLE = "m2 model indirect - Lab 3", 
  VARIABLE = 
   "usevar =
    totexpk classk        
    tmathssk, treadssk
    tchXclas;   ", 
  
  DEFINE = 
    "center totexpk (grandmean);
     tchXclas = totexpk*classk;  ! create interaction term" ,
  
  ANALYSIS = 
    "estimator = mlr; ",
  
  MODEL = 
   "treadssk on classk totexpk tchXclas;
    
    [tmathssk](b0);
    
    tmathssk on
    classk (b1)
    totexpk (b2)
    tchXclas (b3); ",
  
  MODELCONSTRAINT =  
  "LOOP(x,-1,1,0.01);
   PLOT(small regular);
   new(hi_small lo_small hi_regular lo_regular diff_hi); 
     small = b0 + b2*x;
   regular = b0 + b1 + (b2+b3)*x;
   
     hi_small = b0 + b2*(9.3);
     lo_small = b0 + b2*(-9.3);
   hi_regular = b0 + b1 + (b2 + b3)*(9.3);
   lo_regular = b0 + b1 + (b2 + b3)*(-9.3);
      diff_hi = hi_small - hi_regular; ",
  
  OUTPUT = "sampstat standardized modindices (3.84)",
  
  PLOT = "type=plot3;",
  
  usevariables = colnames(mod_data),   
  rdata = mod_data)                    

m2_model_fit <- mplusModeler(m2_model,
                    dataout=here("mplus_files", "caschools.dat"),       
                    modelout=here("mplus_files", "model2_caschools.inp"),
                    check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

## Create the simple slope plot from Mplus model output 

______________________________________________

```{r, eval=FALSE}

 simp_slope2 <- data.frame(m2_model_fit[["results"]][["parameters"]][["unstandardized"]]) %>% 
   filter(paramHeader == "New.Additional.Parameters")  %>%
   filter(param!= "DIFF_HI") %>% 
   select(param, est, se) %>% 
   mutate(size = case_when(
      param %in% c("HI_SMALL", "LO_SMALL") ~ "Small",
      param %in% c("HI_REGUL", "LO_REGUL") ~ "Regular")) %>% 
   mutate(experience = case_when(
      param %in% c("HI_SMALL", "HI_REGUL") ~ 9.3,
      param %in% c("LO_SMALL", "LO_REGUL") ~ -9.3)) 
   
```

```{r, eval=FALSE}
# un-center 'experience' so values on x-axis are on the original scale
mean_exp <- mean(mod_data$totexpk)
plot_data2 <- simp_slope2 %>% mutate(experience = experience + mean_exp)

ggplot(plot_data2, aes(x=experience, y=est, color=size, group=size)) +
  geom_point(size=0) +
  geom_line() +
  geom_errorbar(aes(ymin=est-se, ymax=est+se), width=.25) +
  scale_x_continuous( breaks = c(seq(0,18,2))) +
  labs(title = "Simple Slopes Graph",
       subtitle = "Math test score predicted by years of teaching experience in small & regular classrooms",
           x = "Teaching Experience (years)",
           y = "Math test score") +
  theme_minimal()

ggsave(here("figures", "m2_simple_slope.png"), height = 6, width = 8)
```

```{r, echo=FALSE, eval=TRUE, out.width = "80%", out.height= "80%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/m2_simple_slope.png")
```


______________________________________________

## End 

______________________________________________

## References

Hallquist, M. N., & Wiley, J. F. (2018). MplusAutomation: An R Package for Facilitating Large-Scale Latent Variable Analyses in Mplus. Structural equation modeling: a multidisciplinary journal, 25(4), 621-638.

Horst, A. (2020). Course & Workshop Materials. GitHub Repositories, https://https://allisonhorst.github.io/

Muthén, L.K. and Muthén, B.O. (1998-2017).  Mplus User’s Guide.  Eighth Edition. Los Angeles, CA: Muthén & Muthén

R Core Team (2017). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL http://www.R-project.org/

Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686
