---
title: "Mediation"
author: "Norwegian University of Science and Technology - A Course in `MplusAutomation`"
subtitle: '*Adam Garber*'
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
library(here);library(glue);library(linguisticsdown)
```

______________________________________________

## Lab preparation

______________________________________________

## Load packages 
```{r, eval=TRUE}

library(mediation)
library(tidyverse)
library(MplusAutomation)
library(here)
library(gt)
library(gtsummary)

```

______________________________________________

## Lab outline

1. Estimate a mediation model using the {`mediation`} package
2. Estimate the same model using the Structural Equation Modeling (SEM) framework with {`MplusAutomation`}
3. For the second empirical example, estimate parallel models using the `mediation` and `SEM` methods

______________________________________________

## A quick detour - Equivalent models 

```{r, echo=FALSE, eval=TRUE, out.width = "80%", out.height= "80%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/equiv_models.png")
```

*Figure*. Picture adapted from SEM slides by Sacha Epskamp http://sachaepskamp.com/files/SEM22019/SEM2_2019_Week2_slides.pdf

______________________________________________

## The empirical examples of mediation used in this exercise are from the following article

Tingley, D., Yamamoto, T., Hirose, K., Keele, L., & Imai, K. (2014). Mediation: R package for causal mediation analysis.

https://cran.r-project.org/web/packages/mediation/vignettes/mediation.pdf

______________________________________________

## Data source for example 1

Brader T, Valentino NA, Suhat E (2008). **What Triggers Public Opposition to Immigration?
Anxiety, Group Cues, and Immigration.** American Journal of Political Science, 52(4),
959–978.

https://onlinelibrary.wiley.com/doi/pdf/10.1111/j.1540-5907.2008.00353.x

To see metadata run - `?framing` 

______________________________________________

Read in the `framing` dataset 
```{r}

set.seed(4212020)

data("framing", package = "mediation")

framing <- droplevels(framing) %>%  # drop factor levels with frequency zero
  mutate(emo = emo - 2)


```

______________________________________________

Take a look at variables used in the mediation model 
```{r, eval=TRUE, echo=FALSE}

var_table <- tribble(
   ~"Name",      ~"Labels",  ~"Variable status",                                   
 #--------------|--------------------------------|--------------------------------|,
  "emo"     , "Measure of subjects' negative feeling during the experiment (1-10). 1 indicates the most negative feeling." , "Mediator (Z)"   , 
  "treat"   , "Framing codition interaction term. News story with conditions tone (Negative/Positive) and ethnic identity cue (Latino/European)", "Treatment (X)", 
  "cong_mesg", "Whether subjects requested sending an anti-immigration message to Congress on their behalf." , "Outcome (Y)", 
  "age"  , "Age of subject (18-85)"        , "pre-treatment covariate"          ,            
  "educ"  , "Education (1-4)"        , "pre-treatment covariate"         ,
  "gender"  , "Gender (Male/Female)"      , "pre-treatment covariate"       ,
  "income"  , "Subjects' income, measured as a 19-point scale."        , "pre-treatment covariate"       )

var_table %>% gt()
```

______________________________________________

Look at descriptives table for the framing dataset using {`gtsummary`} 
```{r}

table_data <- framing %>% 
  select(emo, treat, cong_mesg, age, educ, gender, income)

table1 <- tbl_summary(table_data,
                      statistic = list(all_continuous() ~ "{mean} ({sd})"),
                      missing = "no") %>%
  bold_labels() 

table1
```

______________________________________________

```{r, echo=FALSE, eval=TRUE, out.width = "80%", out.height= "80%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/m1_framing_med.png")
```

______________________________________________

## Estimate a mediation model in R using {`mediation`} 

step 1: fit a linear model of the mediator (`emo`) regressed on treatment (`treat`) and pre-treatment covariates
```{r}

med_fit <- lm(emo ~ treat + age + educ + gender + income, 
              data = framing)

```

______________________________________________

step 2: fit a general linear model (`glm`) with the binary outcome variable `cong_mesg` regressed on treatment (`treat`), mediator, and pre-treatment covariates
```{r}

out_fit <- glm(cong_mesg ~ emo + treat + age + educ + gender + income,
               data = framing,
               family = binomial("probit"))
```

______________________________________________

step 3: estimate the mediation effects with bias corrected bootstrapped confidence intervals
```{r}

med_out <- mediate(med_fit, out_fit, treat = "treat", mediator = "emo",
                   boot = TRUE, boot.ci.type ="bca", sims = 100)

summary(med_out)

```

______________________________________________

## Run mediation model 1 using the Structural Equation Modeling framework with {`MplusAutomation`}

______________________________________________

```{r, eval=FALSE}

m1_mediate  <- mplusObject(
  TITLE = "m1 mediate framing", 
  VARIABLE = 
   "usevar =
    cong_mesg emo treat age
    educ gender income;
  
    categorical = cong_mesg; ! outcome is binary",            
  
  ANALYSIS =  "bootstrap = 500; ! set number of bootstrap samples (500 for example purposes)" , 
  
  MODEL = 
   "emo on treat age educ gender income;            ! mediator linear regression 
    cong_mesg on emo treat age educ gender income;  ! outcome GLM regression
    
    Model indirect:
    cong_mesg ind treat;" ,
  
  OUTPUT = 
    "sampstat standardized cinterval (bcbootstrap); ! bias-corrected bootstrap", 
  
  PLOT = "type=plot2;",
  
  usevariables = colnames(framing),   
  rdata = framing)                    

m1_mediate_fit <- mplusModeler(m1_mediate,
                    dataout=here("15-mediation", "mplus_files", "framing.dat"),       
                    modelout=here("15-mediation", "mplus_files", "m1_mediate_framing.inp"),
                    check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

**Model 1 Mplus output**

    STANDARDIZED TOTAL, TOTAL INDIRECT, SPECIFIC INDIRECT, AND DIRECT EFFECTS

                        Estimate       S.E.  Est./S.E.    P-Value

    Effects from TREAT to CONG_MES
    
      Total                0.109      0.075      1.453      0.146
      Total indirect       0.101      0.031      3.253      0.001
    
      Specific indirect 1
        CONG_MES
        EMO
        TREAT              0.101      0.031      3.253      0.001
    
      Direct
        CONG_MES
        TREAT              0.008      0.071      0.119      0.905
    
______________________________________________

## Data source for example 2

Vinokur AD, Price RH, Schul Y (1995). **Impact of the JOBS Intervention on Unemployed Workers Varying in Risk for  Depression.** American Journal of Community Psychology, 23(1), 39–74.

https://link.springer.com/content/pdf/10.1007/BF02506922.pdf

To see metadata run - `?jobs` 

______________________________________________

*Note:* For this example we will ignore the issue of non-compliance addressed in Tingley et al. (2014) as this causal inference topic is beyond the scope of this course. 

Read in the data from the job search intervention study (`jobs`)
```{r}

data("jobs", package = "mediation")

```

Take a look at variables used in the mediation model 
```{r, echo=FALSE}
jobs_vars <- tibble::tribble(
     ~Name, ~Label, 
"depress2 (Y)","Measure of depressive symptoms post-treatment.",
"treat (X)","Indicator variable for whether participant was randomly selected for the JOBS II training program. 1 = assignment to participation.", 
"job_dich (Z)","The job_seek measure recoded into two categories of high and low. 1 = high job search self-efficacy.",
"sex", "Indicator variable for sex. 1 = female",                                                                  "age","Age in years.",
"marital","Factor with five categories for marital status.",                                                      "nonwhite", "Indicator variable for race. 1 = nonwhite.",
"educ","Factor with five categories for educational attainment.",
"income","Factor with five categories for level of income.",
  )

jobs_vars %>% gt()
```

______________________________________________

Look at descriptives of the framing dataset using {`gtsummary`} 
```{r, eval=FALSE}

jobs_desc <- jobs %>% 
  dplyr::select(depress2, job_dich, treat, sex, age, marital, nonwhite, educ, income)

tablej <- tbl_summary(jobs_desc,
                      statistic = list(all_continuous() ~ "{mean} ({sd})"),
                      missing = "no" ) %>%
  bold_labels() 

tablej
```


______________________________________________

```{r, echo=FALSE, eval=TRUE, out.width = "80%", out.height= "80%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/m2_jobs_med.png")
```

______________________________________________

step 1: fit a binomial logist model using `glm` with the binary mediator (`job_dich`) regressed on treatment (`treat`) and pre-treatment covariates
```{r}

jmed_fit <- glm(job_dich ~ treat + sex + age + marital +
                   nonwhite + educ + income,
                 data = jobs, family = binomial)
```

______________________________________________

step 2: fit a linear model with depression score (`depress2`) regressed on treatment, mediator, and pre-treatment covariates
```{r}
jout_fit <- lm(depress2 ~ job_dich + treat +
                 sex + age + marital + nonwhite + educ + income,
               data = jobs)

```

______________________________________________

step 3: Estimate the mediation effects with bias corrected bootstrapped confidence intervals.
```{r}

jmed_out <- mediate(jmed_fit, jout_fit, treat = "treat", mediator = "job_dich",
                   boot = TRUE, boot.ci.type ="bca", sims = 100)

summary(jmed_out)

```

______________________________________________

## Run mediation model 2 as a SEM model with {`MplusAutomation`}

______________________________________________


```{r, eval=FALSE}

m2_jmediate  <- mplusObject(
  
  TITLE = "m2 jobs mediate", 
  
  VARIABLE = 
   "usevar = treat sex
    age marital nonwhite
    educ income depress2 job_dich;
                 
    categorical = job_dich; ! moderator is binary",            
  
  ANALYSIS =  
   "bootstrap = 500; ! set number of bootstrap draws (500 for example purposes)" , 
  
  MODEL = 
   "job_dich on treat sex age marital nonwhite educ income;
   
    depress2 on job_dich treat sex age marital nonwhite educ income;  
    
    Model indirect:
    depress2 ind treat;" ,
  
  OUTPUT = 
    "sampstat standardized cinterval (bcbootstrap); ! bias-corrected bootstrap", 
  
  PLOT = "type=plot2;",
  
  usevariables = colnames(jobs),   
  rdata = jobs)                    

m2_jmediate_fit <- mplusModeler(m2_jmediate,
                    dataout=here("15-mediation", "mplus_files", "jobs.dat"),       
                    modelout=here("15-mediation", "mplus_files", "m2_mediate_jobs.inp"),
                    check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

**Model 2 Mplus output**

STANDARDIZED TOTAL, TOTAL INDIRECT, SPECIFIC INDIRECT, AND DIRECT EFFECTS

                        Estimate       S.E.  Est./S.E.    P-Value

    Effects from TREAT to DEPRESS2
    
      Total               -0.043      0.033     -1.306      0.192
      Total indirect      -0.029      0.012     -2.385      0.017
    
      Specific indirect 1
        DEPRESS2
        JOB_DICH
        TREAT             -0.029      0.012     -2.385      0.017
    
      Direct
        DEPRESS2
        TREAT             -0.015      0.033     -0.440      0.660

______________________________________________

## Run model 3 including the mediator*treatement interaction (potential outcomes framework)

For further reading on this topic see chapter 3 of *Regression and mediation analysis using Mplus* (Muthen et al., 2017) 

```{r, echo=FALSE, eval=TRUE, out.width = "80%", out.height= "80%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/MX_mediate_lab4.png")
```

______________________________________________

```{r, eval=FALSE}

m3_jmed  <- mplusObject(
  
  TITLE = "m3 MX jobs mediate", 
  
  VARIABLE = 
   "usevar =
    treat sex age marital nonwhite
    educ income depress2 job_dich mx; ", 
  
  DEFINE = "mx = job_dich*treat;",
  
  ANALYSIS =  "bootstrap = 500; ", 
  
  MODEL = 
   "job_dich on treat sex age marital nonwhite educ income;           
    depress2 on job_dich treat mx sex age marital nonwhite educ income;  
    
    Model indirect:
    depress2 MOD job_dich mx treat; ",
  
  OUTPUT = 
    "sampstat cinterval(bootstrap); ", 
  
  usevariables = colnames(jobs),   
  rdata = jobs)                    

m3_jmed_fit <- mplusModeler(m3_jmed,
                    dataout=here("15-mediation", "mplus_files", "jobs.dat"),       
                    modelout=here("15-mediation", "mplus_files", "m3_mediate_jobs.inp"),
                    check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

**Model 3 Mplus output**

TOTAL, INDIRECT, AND DIRECT EFFECTS BASED ON COUNTERFACTUALS (CAUSALLY-DEFINED EFFECTS)

    Effects from TREAT to DEPRESS2

                        Estimate       S.E.  Est./S.E.    P-Value
    
      Tot natural IE      -0.026      0.011     -2.357      0.018
      Pure natural DE     -0.022      0.055     -0.401      0.688
      Total effect        -0.048      0.055     -0.878      0.380
    
     Other effects
    
      Pure natural IE     -0.023      0.012     -1.938      0.053
      Tot natural DE      -0.026      0.052     -0.494      0.621
      Total effect        -0.048      0.055     -0.878      0.380

______________________________________________

## References

Brader T, Valentino NA, Suhat E (2008). What Triggers Public Opposition to Immigration?
Anxiety, Group Cues, and Immigration. American Journal of Political Science, 52(4),
959–978.

Hallquist, M. N., & Wiley, J. F. (2018). MplusAutomation: An R Package for Facilitating Large-Scale Latent Variable Analyses in Mplus. Structural equation modeling: a multidisciplinary journal, 25(4), 621-638.

Ingels, S. J., Pratt, D. J., Herget, D. R., Burns, L. J., Dever, J. A., Ottem, R., ... & Leinwand, S. (2011). High School Longitudinal Study of 2009 (HSLS: 09): Base-Year Data File Documentation. NCES 2011-328. National Center for Education Statistics.

Muthén, B. O., Muthén, L. K., & Asparouhov, T. (2017). Regression and mediation analysis using Mplus. Los Angeles, CA: Muthén & Muthén.

Muthén, L.K. and Muthén, B.O. (1998-2017).  Mplus User’s Guide.  Eighth Edition. Los Angeles, CA: Muthén & Muthén

R Core Team (2017). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL http://www.R-project.org/

Tingley, D., Yamamoto, T., Hirose, K., Keele, L., & Imai, K. (2014). Mediation: R package for causal mediation analysis.

Vinokur AD, Price RH, Schul Y (1995). Impact of the JOBS Intervention on Unemployed Workers Varying in Risk for  Depression. American Journal of Community Psychology, 23(1), 39–74.

Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686