---
title: "Mixture Models with Covariates & Distal Outcomes"
author: "Norwegian University of Science and Technology - A Course in `MplusAutomation`"
subtitle: '*Adam Garber*'
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
library(here);library(glue);library(linguisticsdown)
```

______________________________________________

## Preparation

______________________________________________

## Change starting location to folder `20-three-step`
```{r}

source("rep_functions.R")

change_here(glue("{project_location}/20-three-step"))

here()

```

______________________________________________

## Data source:

1. The first example utilizes a dataset on undergraduate *Cheating* available from the `poLCA` package (Dayton, 1998): [$\color{blue}{\text{See documentation here}}$](https://cran.r-project.org/web/packages/poLCA/poLCA.pdf)

2. The second examples utilizes the public-use dataset, *The Longitudinal Survey of American Youth* (**LSAY**): [$\color{blue}{\text{See documentation here}}$](https://www.lsay.org/)

______________________________________________

Load packages
```{r, eval=TRUE}

library(naniar)
library(tidyverse)
library(haven)
library(glue)
library(MplusAutomation)
library(here)
library(janitor)
library(gt)
library(poLCA)

```

____________________________________

Incorporating distal outcome variables with mixture models 

**Note**: Prior to adding covariates or distals enumeration must be conducted. 

[$\color{blue}{\text{See Lab 7 for examples of enumeration with MplusAutomation}}$](https://garberadamc.github.io/project-site/Lab8-Intro-mixture)

____________________________________

## `DU3step` auxiliary variable integration

- Using the `DU3step` you can specify distal relations but cannot specify models with covariates & distals

____________________________________

## Application: Undergraduate Cheating behavior

"Dichotomous self-report responses by 319 undergraduates to four questions about cheating behavior" (poLCA, 2016).

____________________________________


Prepare data 
```{r, eval=TRUE}

data(cheating)

cheating <- cheating %>% clean_names() 

df_cheat <-  cheating %>%                                 
  dplyr::select(1:4) %>%                                  
  mutate_all(funs(.-1)) %>%                               
  mutate(gpa = cheating$gpa)
```

Run the **DU3step** model with `gpa` as distal outcome
```{r, eval=FALSE}

m_stepdu  <- mplusObject(
  TITLE = "DU3STEP add distal GPA", 
  VARIABLE = 
   "categorical = lieexam-copyexam; 
    usevar = lieexam-copyexam;
    auxiliary = gpa (du3step);
    classes = c(2);",
  
  ANALYSIS = 
   "estimator = mlr; 
    type = mixture;
    starts = 500 100; 
    processors = 10;",
  
  OUTPUT = "sampstat patterns tech11 tech14;",
  
  PLOT = 
    "type = plot3; 
     series = lieexam-copyexam(*);",
  
  usevariables = colnames(df_cheat),
  rdata = df_cheat)

m_stepdu_fit <- mplusModeler(m_stepdu, 
                dataout=here("du3step_mplus", "lca_du3step.dat"),
                modelout=here("du3step_mplus", "c2_lca_du3step.inp") ,
                check=TRUE, run = TRUE, hashfilename = FALSE)

```

______________________________________________

## Application: Longitudinal Study of American Youth, **Science Attitudes** 

______________________________________________

Load data
```{r, eval=TRUE}

lsay_data <- read_csv("https://garberadamc.github.io/project-site/data/lca_lsay_sci.csv",
  na = c("9999", "9999.00")) %>%               
  clean_names() %>%                                                                                 
  dplyr::select(1:5, female, mathg12,                                                               
         Enjoy = ab39m, Useful = ab39t,                                                             
         Logical = ab39u, Job = ab39w, Adult = ab39x)                                               

```

Use {`naniar`} to look at missing on covariates and distals

```{r, eval=TRUE}

naniar::gg_miss_var(lsay_data)

```


____________________________________

## **Manual 3-step** 

- Adding covariates and distals to a mixture model
- Often called "*auxiliary variable integration*"

____________________________________

## Step 1

____________________________________

```{r, eval=FALSE}

step1  <- mplusObject(
  TITLE = "Step1 - 3step LSAY", 
  VARIABLE = 
  "categorical = Enjoy-Adult; 
   usevar = Enjoy-Adult;
    
   classes = c(4); 
    
   auxiliary =   ! list all potential covariates and distals here
   female        ! covariate
   mathg12;      ! distal math test score in 12th grade ",
  
  ANALYSIS = 
   "estimator = mlr; 
    type = mixture;
    starts = 500 100;",
  
  SAVEDATA = 
   "File=3step_savedata.dat;
    Save=cprob;
    Missflag= 999;",
  
  OUTPUT = "sampstat residual tech11 tech14",
  
  PLOT = 
    "type = plot3; 
    series = Enjoy-Adult(*);",
  
  usevariables = colnames(lsay_data),
  rdata = lsay_data)

step1_fit <- mplusModeler(step1,
                            dataout=here("3step_mplus", "Step1_3step_LSAY.dat"),
                            modelout=here("3step_mplus", "Step1_3step_LSAY.inp") ,
                            check=TRUE, run = TRUE, hashfilename = FALSE)
```

____________________________________

## Step 2 

____________________________________

Extract logits for the classification probabilities for the most likely latent class 
```{r, eval=FALSE}

logit_cprobs <- as.data.frame(step1_fit[["results"]]
                                       [["class_counts"]]
                                       [["logitProbs.mostLikely"]])
```

Extract saved dataset which is part of the mplusObject "step1_10_fit"
```{r, eval=FALSE}

savedata <- as.data.frame(step1_fit[["results"]]
                                   [["savedata"]])
```

Rename the column in savedata named "C" and change to "N"
```{r, eval=FALSE}

colnames(savedata)[colnames(savedata)=="C"] <- "N"

```


Run step 2 
```{r, eval=FALSE}
step2  <- mplusObject(
  TITLE = "Step2 - 3step LSAY", 
  
  VARIABLE = 
 "nominal=N;
  USEVAR = n;
  missing are all (999); 
  classes = c(4); ",
  
  ANALYSIS = 
 "estimator = mlr; 
  type = mixture; 
  starts = 0;",
  
  MODEL = 
    glue(
 "%C#1%
  [n#1@{logit_cprobs[1,1]}];
  [n#2@{logit_cprobs[1,2]}];
  [n#3@{logit_cprobs[1,3]}];
  
  %C#2%
  [n#1@{logit_cprobs[2,1]}];
  [n#2@{logit_cprobs[2,2]}];
  [n#3@{logit_cprobs[2,3]}];
  
  %C#3%
  [n#1@{logit_cprobs[3,1]}];
  [n#2@{logit_cprobs[3,2]}];
  [n#3@{logit_cprobs[3,3]}];
   
  %C#4%
  [n#1@{logit_cprobs[4,1]}];
  [n#2@{logit_cprobs[4,2]}];
  [n#3@{logit_cprobs[4,3]}];"),
 
  usevariables = colnames(savedata), 
  rdata = savedata)

step2_fit <- mplusModeler(step2, 
                            dataout=here("3step_mplus", "Step2_3step_LSAY.dat"), 
                            modelout=here("3step_mplus", "Step2_3step_LSAY.inp"), 
                            check=TRUE, run = TRUE, hashfilename = FALSE)
```

____________________________________

## Step 3 

____________________________________

Model with 1 covariate and 1 distal outcome
```{r, eval=FALSE}
step3  <- mplusObject(
  TITLE = "Step3 - 3step LSAY", 
  
  VARIABLE = 
 "nominal=N;
  usevar = n;
  missing are all (999);
  classes = c(4);
  
  usevar = female mathg12;" ,
  
  ANALYSIS = 
 "estimator = mlr; 
  type = mixture; 
  starts = 0;",
  
  MODEL =
  glue(
 " %OVERALL%
  
  C on female;      ! covariate as predictor of C

     %C#1%
  [n#1@{logit_cprobs[1,1]}];
  [n#2@{logit_cprobs[1,2]}];
  [n#3@{logit_cprobs[1,3]}];
  
  [mathg12](m1);    ! conditional distal mean 
  mathg12;          ! conditional distal variance (freely estimated)

  %C#2%
  [n#1@{logit_cprobs[2,1]}];
  [n#2@{logit_cprobs[2,2]}];
  [n#3@{logit_cprobs[2,3]}];
  
  [mathg12](m2);
  mathg12;
  
  %C#3%
  [n#1@{logit_cprobs[3,1]}];
  [n#2@{logit_cprobs[3,2]}];
  [n#3@{logit_cprobs[3,3]}];
  
  [mathg12](m3);
  mathg12;

  %C#4%
  [n#1@{logit_cprobs[4,1]}];
  [n#2@{logit_cprobs[4,2]}];
  [n#3@{logit_cprobs[4,3]}];
  
  [mathg12](m4);
  mathg12; "),
  
  MODELCONSTRAINT = 
   "New (diff12 diff13 diff23 
    diff14 diff24 diff34);
  
    diff12 = m1-m2;  ! test pairwise distal mean differences
    diff13 = m1-m3;
    diff23 = m2-m3;
    diff14 = m1-m4;
    diff24 = m2-m4;
    diff34 = m3-m4;",
  
  MODELTEST = "     ! omnibus test of distal means
    m1=m2;
    m2=m3;
    m3=m4;",
 
  usevariables = colnames(savedata), 
  rdata = savedata)

step3_fit <- mplusModeler(step3,
               dataout=here("3step_mplus", "Step3_3step_LSAY.dat"), 
               modelout=here("3step_mplus", "Step3_3step_LSAY.inp"), 
               check=TRUE, run = TRUE, hashfilename = FALSE)
```

**End of manual 3-step**

____________________________________

## Model wtih latent categorical variable ($C_k$) as moderator  

____________________________________


```{r, eval=FALSE}
step3mod  <- mplusObject(
  TITLE = "Step3 - 3step LSAY", 
  
  VARIABLE = 
 "nominal=N;
  usevar = n;
  missing are all (999);
  classes = c(4);
  
  usevar = female mathg12;" ,
  
  ANALYSIS = 
 "estimator = mlr; 
  type = mixture; 
  starts = 0;",
  
  MODEL =
  glue(
 "!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
  !!!!!DISTAL = mathg12, COVARIATE = female, MODERATOR = C!!!!!
  !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
  %OVERALL%
  mathg12 on female;
  mathg12;

     %C#1%
  [n#1@{logit_cprobs[1,1]}];
  [n#2@{logit_cprobs[1,2]}];
  [n#3@{logit_cprobs[1,3]}];
  
  mathg12 on female(s1);  ! conditional slope (class 1)
  [mathg12](m1);          ! conditional distal mean
  mathg12;                ! conditional distal variance (freely estimated)

  %C#2%
  [n#1@{logit_cprobs[2,1]}];
  [n#2@{logit_cprobs[2,2]}];
  [n#3@{logit_cprobs[2,3]}];
  
  mathg12 on female(s2);
  [mathg12](m2);
  mathg12;
  
  %C#3%
  [n#1@{logit_cprobs[3,1]}];
  [n#2@{logit_cprobs[3,2]}];
  [n#3@{logit_cprobs[3,3]}];
  
  mathg12 on female(s3);
  [mathg12](m3);
  mathg12;

  %C#4%
  [n#1@{logit_cprobs[4,1]}];
  [n#2@{logit_cprobs[4,2]}];
  [n#3@{logit_cprobs[4,3]}];
  
  mathg12 on female(s4);
  [mathg12](m4);
  mathg12; "),
  
  MODELCONSTRAINT = 
   "New (slope12 slope13 slope23 
    slope14 slope24 slope34);
  
    slope12 = s1-s2;  ! test pairwise slope differences
    slope13 = s1-s3;
    slope23 = s2-s3;
    slope14 = s1-s4;
    slope24 = s2-s4;
    slope34 = s3-s4;",
  
  MODELTEST = " ! can run only a single Omnibus test per model 
    s1=s2;
    s2=s3;
    s3=s4;",
 
  usevariables = colnames(savedata), 
  rdata = savedata)

step3mod_fit <- mplusModeler(step3mod,
               dataout=here("3step_mplus", "Step3_moderation_LSAY.dat"), 
               modelout=here("3step_mplus", "Step3_moderation_LSAY.inp"), 
               check=TRUE, run = TRUE, hashfilename = FALSE)
```



## References

Drew A. Linzer, Jeffrey B. Lewis (2011). poLCA: An R Package for Polytomous Variable Latent Class Analysis. Journal of Statistical Software, 42(10), 1-29. URL http://www.jstatsoft.org/v42/i10/.

Hallquist, M. N., & Wiley, J. F. (2018). MplusAutomation: An R Package for Facilitating Large-Scale Latent Variable Analyses in Mplus. Structural equation modeling: a multidisciplinary journal, 25(4), 621-638.

Miller, J. D., Hoffer, T., Suchner, R., Brown, K., & Nelson, C. (1992). LSAY codebook. Northern Illinois University.

Muthén, B. O., Muthén, L. K., & Asparouhov, T. (2017). Regression and mediation analysis using Mplus. Los Angeles, CA: Muthén & Muthén.

Muthén, L.K. and Muthén, B.O. (1998-2017).  Mplus User’s Guide.  Eighth Edition. Los Angeles, CA: Muthén & Muthén

R Core Team (2017). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL http://www.R-project.org/

Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686

