---
title: "Observed Response Patterns in Latent Class Analysis"
author: "Norwegian University of Science and Technology - A Course in `MplusAutomation`"
subtitle: '*Adam Garber*'
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
library(here);library(glue);library(linguisticsdown)
```

***

## Lab preparation

***

## Data source: Longitudinal Study of American Youth, **Science Attitudes** 

[$\color{blue}{\text{See documentation about the LSAY here.}}$](https://www.lsay.org/)

***

Load packages
```{r, eval=TRUE}

library(tidyverse)
library(glue)
library(MplusAutomation)
library(here)
library(janitor)
library(gt)
library(DT)
library(plotly)
library(gg3D)
library(gganimate)
library(viridis)
library(hrbrthemes)

```

***

## Exploring observed response patterns

***

Load data
```{r, eval=TRUE}

lsay_data <- read_csv("https://garberadamc.github.io/project-site/data/lca_lsay_sci.csv", 
                      na = c("9999", "9999.00")) %>%                                     
  clean_names() %>%                                                                      
  dplyr::select(1:5, Enjoy = ab39m, Useful = ab39t,                                      
                     Logical = ab39u, Job = ab39w, Adult = ab39x)                        


```

Use {`DT::datatable()`} to take a look at the data
```{r, eval=TRUE}

datatable(lsay_data, rownames = FALSE, filter="top",
          options = list(pageLength = 5, scrollX=T) )
```

```{r, echo=FALSE, eval=FALSE, out.width = "65%", out.height= "65%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/LCA_Diagram_SEMLAB8.png")
```

*Figure*. Path diagram of science attitude indicators.

***

Save response frequencies for the 4 class model with `response is _____.dat`. 
```{r, eval=FALSE}

patterns  <- mplusObject(
  TITLE = "C4 LCA - Save response patterns", 
  VARIABLE = 
  "categorical = Enjoy-Adult; 
   usevar = Enjoy-Adult;
    
   classes = c(4);",
  
  ANALYSIS = 
   "estimator = mlr; 
    type = mixture;
    starts = 500 100;",
  
  SAVEDATA = 
   "File=3step_savedata.dat;
    Save=cprob;
    Missflag= 999;
    !!!!!!!! Code to save response frequency data !!!!!!!!
    response is resp_patterns.dat; 
    !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!",
  
  OUTPUT = "sampstat residual patterns tech10 tech11 tech14",
  
  PLOT = 
    "type = plot3; 
    series = Enjoy-Adult(*);",
  
  usevariables = colnames(lsay_data),
  rdata = lsay_data)

patterns_fit <- mplusModeler(patterns,
                dataout=here("21-response-patterns", "resp_pattrn", "LSAY.dat"),
                modelout=here("21-response-patterns", "resp_pattrn", "patterns.inp") ,
                check=TRUE, run = TRUE, hashfilename = FALSE)
```

Read in observed respnse pattern data
```{r}

patterns <- read_table2(here("21-response-patterns", "resp_pattrn", "resp_patterns.dat"),
                        col_names=FALSE, na = "*")                                                   

colnames(patterns) <- c("Frequency", "ENJOY", "USEFUL", "LOGICAL", "JOB", "ADULT",                   
                      "CPROB1", "CPROB2", "CPROB3", "CPROB4", "C_MODAL")                             

```


Order responses by highest frequency 
```{r,eval=TRUE}

order_highest <- patterns %>% 
  arrange(desc(Frequency)) 

```

```{r,eval=TRUE}

loop_cond  <- lapply(1:4, function(k) {       
order_cond <- patterns %>%                    
  filter(C_MODAL == k) %>%                    
  arrange(desc(Frequency)) %>%                
  head(5)                                     
  })                                          

table_data1 <- bind_rows(loop_cond) %>%       
  as.data.frame()                             

table_data2 <-  rbind(order_highest[1:5,], table_data1) 
```

Use `{gt}` to make a nicely formatted table
```{r,eval=TRUE}

table_data2 %>% 
  gt() %>%
    tab_header(
    title = md("**Observed response patterns, estimated frequencies, estimated posterior
                  class probabilities, and modal class assignment.**"),
    subtitle = md("&nbsp;")) %>% 
    tab_source_note(
    source_note = md("Data Source: **Longitudinal Study of American Youth.**")) %>%
    cols_label(
    ENJOY = "Enjoy",
    USEFUL = "Useful",
    LOGICAL = "Logical",
    JOB = "Job",
    ADULT = "Adult",
    CPROB1 = html("P<sub>k=1"),
    CPROB2 = html("P<sub>k=2"),
    CPROB3 = html("P<sub>k=3"),
    CPROB4 = html("P<sub>k=4"),
    C_MODAL = md("*k*")) %>% 
  tab_row_group(
    group = "Unconditional response patterns ordered by highest frequency",
    rows = 1:5) %>%
  tab_row_group(
    group = "k=1 conditional response pattern ordered by highest frequency",
    rows = 6:10) %>%
  tab_row_group(
    group = "k=2 conditional response pattern ordered by highest frequency",
    rows = 11:15)  %>%
  tab_row_group(
    group = "k=3 conditional response pattern ordered by highest frequency",
    rows = 16:20) %>%
  tab_row_group(
    group = "k=4 conditional response pattern ordered by highest frequency",
    rows = 21:25) %>%
    row_group_order(
      groups = c("Unconditional response patterns ordered by highest frequency",
                 "k=1 conditional response pattern ordered by highest frequency",
                 "k=2 conditional response pattern ordered by highest frequency",
                 "k=3 conditional response pattern ordered by highest frequency",
                 "k=4 conditional response pattern ordered by highest frequency")) %>% 
  tab_options(column_labels.font.weight = "bold")


```

***

## Visualizing observed response patterns

***

Order rows by modal assignment (*K*)
```{r, eval=TRUE}

order_modal <- patterns %>% 
  arrange(desc(C_MODAL)) %>%
  rownames_to_column() %>% 
  rename('pat_num' = "rowname") %>%
  drop_na(ENJOY:ADULT)

```

Prepare plot data
```{r, eval=TRUE}

p1_long <- order_modal %>% 
  dplyr::select(pat_num:ADULT, C_MODAL) %>% 
  pivot_longer(`ENJOY`:`ADULT`, # The columns I'm gathering together
               names_to = "var", # new column name for existing names
               values_to = "value") %>%  # new column name to store values
  mutate(obs = rep(1:32, each =5)) %>%
  mutate(Class = factor(C_MODAL)) %>% 
  mutate(var = ordered(var,
                      levels = c("ENJOY","USEFUL","LOGICAL","JOB","ADULT"))) %>%
  select(-pat_num, -C_MODAL)

# must first run LCA enumeration (code is out of sequential order)
out_c4 <- readModels(here("21-response-patterns", "resp_pattrn"), filefilter = "patterns", quiet = TRUE)

# extract posterior probabilities 
probs_c4 <- as.data.frame(
  out_c4[["gh5"]][["means_and_variances_data"]]
  [["estimated_probs"]][["values"]]                      
  [seq(2, 10, 2),]) 

rownames(probs_c4) <- c("ENJOY","USEFUL","LOGICAL","JOB","ADULT")

long_c4 <- probs_c4 %>% rownames_to_column() %>%
  rename('var' = "rowname") %>%
  pivot_longer(`V1`:`V4`, # The columns I'm gathering together
               names_to = "c", # new column name for existing names
               values_to = "value") %>% # new column name to store values
  mutate(Class = rep(1:4,5)) %>%
  arrange(Class) %>% 
  mutate(obs = rep(33:36,each=5)) %>%
  mutate(Frequency = rep(c(829,782,619,833),each=5)) %>%
  mutate(var = ordered(var,
                      levels = c("ENJOY","USEFUL","LOGICAL","JOB","ADULT"))) %>%
  select(6,1,3,5,4)

p2_long <- rbind(p1_long, long_c4) %>% 
  mutate(Class = as.numeric(Class))
 
```


Visualize observed response patterns with {`plotly`}
```{r, eval=TRUE}
gg <- ggplot(p2_long, aes(x=var, y=value, color = Class, size=Frequency)) +
  geom_line(aes(as.numeric(var), frame = obs)) +
  scale_color_viridis() + labs(x="Indicator", y= "Probability")

ggplotly(gg) %>%  animation_opts(frame = 1000, transition = 0) %>%
  animation_slider(currentvalue =
                     list(prefix = "Pattern ", font = list(color="red")))

```


Make a 3D plot with packages {`ggplot2`}, {`gg3D`}, and {`gganimate`}.
```{r, eval=FALSE}

theta= 170    # change perspective (tilt)
phi=40        # change perspective (rotation)

resp3d <- ggplot(p1_long, aes(x=as.numeric(var),
                              y=as.numeric(value),
                              z = as.numeric(obs)),
                 alpha = .8) +            
  axes_3D(theta=theta, phi=phi) +
  stat_3D(theta=theta, phi=phi, geom="path",
          aes(colour = Class, size = Frequency), alpha = .8) +
  scale_color_manual(values=c("#FDE725FF", "#DE7065FF", "#238A8DFF", "#482677FF")) +
  theme_void() +
  annotate("text", x = -.3, y = 0.05, label = "Indicators ") +
  annotate("text", x = .35, y = -.4, label = "Probability") +
  annotate("text", x = .25, y = .42, label = "Pattern") +
  annotate("text", x = .2, y = 0, label = "0.0") +
  annotate("text", x = .34, y = -.33, label = "1.0") +
  annotate("text", x = -.05, y = 0, angle = 6,
           label = "Enjoy - Useful - Logical - Job - Adult") +
  transition_states(obs, transition_length=1, state_length=5) +
  shadow_mark(alpha = .1,) +
  labs(title = "Observed response pattern = {closest_state}")


animate(resp3d, fps = 2)
```

```{r, eval=FALSE}
anim_save(here("21-response-patterns", "figures", "responses_3d_anim.gif"), height = 6, width = 8, dpi = "retina")
```

______________________________________________

## Comparing model fit 

Learning objective: Generate a comprehensive model fit summary table.

**Information criteria: model is endorsed by lowest value**:

- `BIC`: $$ =-2*LL+Npar*LN(N) $$
- `aBIC`: $$-2*LL+Npar*LN((N+2)/24)$$
- `CIAC`: $$-2*LL+Npar*(LN(N)+1))$$
- `AWE`: $$ -2*LL+2*Npar*(LN(N)+1.5) $$

______________________________________________

Run a quick enumeration 
```{r, eval=FALSE}

lca_k1_6  <- lapply(1:6, function(k) {
  lca_enum  <- mplusObject(
      
    TITLE = glue("Class {k}"), 
  
    VARIABLE = glue(
    "categorical = Enjoy-Adult; 
     usevar = Enjoy-Adult;
     classes = c({k}); "),
  
  ANALYSIS = 
   "estimator = mlr; 
    type = mixture;
    starts = 200 50; 
    processors = 10;",
  
  OUTPUT = "sampstat residual tech11 tech14;",
  
  PLOT = 
    "type = plot3; 
     series = Enjoy-Adult(*);",
  
  usevariables = colnames(lsay_data),
  rdata = lsay_data)

lca_enum_fit <- mplusModeler(lca_enum, 
    dataout=glue(here("21-response-patterns", "enum_mplus", "lsay.dat")),
    modelout=glue(here("21-response-patterns", "enum_mplus", "c{k}_lca.inp")) ,
    check=TRUE, run = TRUE, hashfilename = FALSE)
})

```

______________________________________________

## Create model fit summary table 

______________________________________________

Extract data and calculate indices derived from the Log Likelihood
```{r}

all_output <- readModels(here("21-response-patterns", "enum_mplus"), quiet = TRUE)

n_size <- all_output[["c1_lca.out"]][["summaries"]][["Observations"]]

enum_extract <- LatexSummaryTable(all_output,                                            
                keepCols=c("Title","Parameters", "LL", "BIC",                            
                           "aBIC", "BLRT_PValue", "T11_VLMR_PValue"),                    
                sortBy = "Title")                                                        
                           
all_fit <- enum_extract %>% 
  mutate(aBIC = -2*LL+Parameters*log((n_size+2)/24)) %>% 
  mutate(CIAC = -2*LL+Parameters*(log(n_size)+1)) %>% 
  mutate(AWE = -2*LL+2*Parameters*(log(n_size)+1.5)) %>%
  mutate(SIC = -.5*BIC) %>% 
  mutate(expSIC = exp(SIC - max(SIC))) %>% 
  mutate(expSUM = sum(expSIC)) %>% 
  mutate(BF = exp(SIC-lead(SIC))) %>% 
  mutate(cmPk = expSIC/expSUM) %>% 
  select(1:5,8:9,7,6,13,14)

```


Format table with package {`gt`}
```{r,eval=TRUE}

all_fit %>% 
  gt() %>%
    tab_header(
    title = md("**Model Fit Summary Table**"), subtitle = md("&nbsp;")) %>% 
    tab_source_note(
    source_note = md("Data Source: **Longitudinal Study of American Youth.**")) %>%
    cols_label(
    Title = "Classes",
    Parameters = md("*NPar*"),
    LL = md("*LL*"),
    T11_VLMR_PValue = html("VLMR"),
    BLRT_PValue = html("BLRT"),
    BF = html("Bayes<br>Factor"),
    cmPk = html("cmP<sub>k")) %>% 
  tab_options(column_labels.font.weight = "bold") %>% 
  fmt_number(10:11,decimals = 2,
             drop_trailing_zeros=TRUE,
             suffixing = TRUE) %>% 
  fmt_number(2:9,decimals = 2)

```


## References

Drew A. Linzer, Jeffrey B. Lewis (2011). poLCA: An R Package for Polytomous Variable Latent Class Analysis. Journal of Statistical Software, 42(10), 1-29. URL http://www.jstatsoft.org/v42/i10/.

Hallquist, M. N., & Wiley, J. F. (2018). MplusAutomation: An R Package for Facilitating Large-Scale Latent Variable Analyses in Mplus. Structural equation modeling: a multidisciplinary journal, 25(4), 621-638.

Miller, J. D., Hoffer, T., Suchner, R., Brown, K., & Nelson, C. (1992). LSAY codebook. Northern Illinois University.

Muthén, B. O., Muthén, L. K., & Asparouhov, T. (2017). Regression and mediation analysis using Mplus. Los Angeles, CA: Muthén & Muthén.

Muthén, L.K. and Muthén, B.O. (1998-2017).  Mplus User’s Guide.  Eighth Edition. Los Angeles, CA: Muthén & Muthén

R Core Team (2017). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL http://www.R-project.org/

Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686