---
title: "Introduction to Latent Profile Analysis"
author: "Norwegian University of Science and Technology - A Course in `MplusAutomation`"
subtitle: '*Adam Garber*'
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
library(here);library(glue);library(linguisticsdown)
```

______________________________________________

```{r, echo=FALSE, eval=TRUE, out.width = "60%", out.height= "60%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/lpa_mix_density.png")
```

*Figure*. Gaussian mixture models. Data simulated from a 2-class model. 

______________________________________________

## Lab preparation

______________________________________________

## Change starting location to folder `22-LPA`
```{r}

source("rep_functions.R")

change_here(glue("{project_location}/22-LPA"))

here()

```

______________________________________________

## Data source:

1. The first example closely follows the vignette used to demonstrate the `tidyLPA` package (Rosenberg, 2019): [$\color{blue}{\text{See detailed documentation of this model here}}$](https://data-edu.github.io/tidyLPA/articles/Introduction_to_tidyLPA.html)

This model utilizes the `PISA` data collected in the U.S. in 2015. To learn more about this data [$\color{blue}{\text{see here}}$](http://www.oecd.org/pisa/data/). 

To access the 2015 US `PISA` data in R use the following code:
`devtools::install_github("jrosen48/pisaUSA15")`
`library(pisaUSA15)`
`open_codebook()`

2. The second examples utilizes 4 test score measures from the public-use dataset, *The Longitudinal Survey of American Youth* (**LSAY**):  [$\color{blue}{\text{See documentation here}}$](https://www.lsay.org/)

______________________________________________

Load packages
```{r, eval=TRUE}

library(naniar)
library(tidyverse)
library(haven)
library(glue)
library(MplusAutomation)
library(here)
library(janitor)
library(gt)
library(tidyLPA)

```

Load data
```{r, eval=TRUE}
pisa <- pisaUSA15
```

___________________________________

## Latent Profile Analysis 

___________________________________

```{r, echo=FALSE, eval=TRUE, out.width = "60%", out.height= "60%", fig.pos="h"}

include_graphics2("https://garberadamc.github.io/project-site/figures/var_covar_matrix.png")

```

*Figure*. Picture adapted from tutorial (Rosenberg, 2019).

- `model 1` Class-invariant / Diagonal: Equal variances, and covariances fixed to 0
- `model 2` Class-varying / Diagonal: Free variances and covariances fixed to 0
- `model 3` Class-invariant / Non-Diagonal: Equal variances and equal covariances
- `model 4` Free variances, and equal covariances
- `model 5` Equal variances, and free covariances 
- `model 6` Class Varying / Non-Diagonal: Free variances and free covariances

___________________________________

Example 1: PISA dataset from the `tidyLPA` package

___________________________________

Enumerate using `estimate_profiles()`:

- Estimate models with classes $K = 1:3$
- Model has 4 continuous indicators
- Default variance-covariance specifications (model 1)
- Add line `scale() %>%` to center indicator means

```{r,eval=FALSE}
lpa_models <- pisa[1:500,] %>%
    select(broad_interest, enjoyment, instrumental_mot, self_efficacy) %>%
    estimate_profiles(1:3,
                      package = "MplusAutomation",
                      ANALYSIS = "starts = 100, 20;",
                      variances = c("equal", "varying"),
                      covariances = c("zero", "varying"))

get_fit(lpa_models)
```

___________________________________

Plot 3-class model

**Note:** single imputation is used in this example as `plot_profiles()` requires complete cases
```{r, eval=TRUE}
pisa[1:200,] %>%
    select(broad_interest, enjoyment, instrumental_mot, self_efficacy) %>%
    single_imputation() %>%
    estimate_profiles(3, package = "MplusAutomation") %>% 
    plot_profiles(sd=FALSE)
```

___________________________________

Plot densities for classes `k = 1:4`
```{r, eval=TRUE}

pisa[1:500, c("broad_interest","enjoyment")] %>%
  single_imputation() %>%
  estimate_profiles(1:4, package = "MplusAutomation") %>%
  plot_density()

```

___________________________________

## Example 2: Math, Science, Physics, and Biology measures (LSAY).

___________________________________


Read in data 
```{r, eval=TRUE}

lsay_data <- read_csv("https://garberadamc.github.io/project-site/data/lsay_lab10.2_lpa.csv")

```

___________________________________

Run a quick enumeration 
```{r, eval=FALSE}

lpa_k14  <- lapply(1:4, function(k) {
  lpa_enum  <- mplusObject(
      
    TITLE = glue("Class {k}"), 
  
    VARIABLE = glue(
    "usevar = mth_scor-bio_scor;
     classes = c({k}); "),
  
  ANALYSIS = 
   "estimator = mlr; 
    type = mixture;
    starts = 200 50; 
    processors = 10;",
  
  OUTPUT = "sampstat residual tech11 tech14;",
  
  PLOT = 
    "type = plot3; 
     series = mth_scor-bio_scor(*);",
  
  usevariables = colnames(lsay_data),
  rdata = lsay_data)

lpa_enum_fit <- mplusModeler(lpa_enum, 
                dataout=glue(here("enum_lpa", "lpa_lsay.dat")),
                modelout=glue(here("enum_lpa", "c{k}_lpa_lsay.inp")) ,
                check=TRUE, run = TRUE, hashfilename = FALSE)
})

```

___________________________________

Plot 3-class profile
```{r, eval=TRUE}
lsay_data[1:500,5:8] %>%
    single_imputation() %>%
    estimate_profiles(1:4, package = "MplusAutomation") %>% 
    plot_profiles(sd=FALSE)
```

*Figure.* Here we see ordered solutions.

___________________________________

Compare model fit.
```{r}
all_output <- readModels(here("enum_lpa"), quiet = TRUE)

enum_extract <- LatexSummaryTable(all_output,                         
                keepCols=c("Title","Parameters", "LL", "BIC",         
                           "aBIC", "BLRT_PValue", "T11_VLMR_PValue"), 
                sortBy = "Title")                                     

enum_extract %>% gt()
```


## References

Hallquist, M. N., & Wiley, J. F. (2018). MplusAutomation: An R Package for Facilitating Large-Scale Latent Variable Analyses in Mplus. Structural equation modeling: a multidisciplinary journal, 25(4), 621-638.

Miller, J. D., Hoffer, T., Suchner, R., Brown, K., & Nelson, C. (1992). LSAY codebook. Northern Illinois University.

Muthén, B. O., Muthén, L. K., & Asparouhov, T. (2017). Regression and mediation analysis using Mplus. Los Angeles, CA: Muthén & Muthén.

Muthén, L.K. and Muthén, B.O. (1998-2017).  Mplus User’s Guide.  Eighth Edition. Los Angeles, CA: Muthén & Muthén

Rosenberg, J. M., van Lissa, C. J., Beymer, P. N., Anderson, D. J., Schell, M. J. & Schmidt, J. A. (2019). tidyLPA: Easily carry out Latent Profile Analysis (LPA) using open-source or commercial software [R package]. https://data-edu.github.io/tidyLPA/

R Core Team (2017). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL http://www.R-project.org/

Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686