---
title: "Template Method - Nested Iterators"
author: "Norwegian University of Science and Technology - A Course in `MplusAutomation`"
subtitle: '*Adam Garber*'
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
library(here);library(glue);library(linguisticsdown)
```

______________________________________________


## Data source:


______________________________________________

## Change starting location to folder `23-template-method`
```{r}

source("rep_functions.R")

change_here(glue("{project_location}/23-template-method"))

here()

```

______________________________________________

Load packages
```{r, eval=TRUE}

library(MplusAutomation)
library(relimp)
library(tidyverse)
library(here)
library(janitor)
library(gt)


```

______________________________________________

## 0. Write the Mplus `template.txt` file 

- This is a special type of Mplus input file that includes the [[/init]] section at the top
- This section of code provides the instructions for doing “iterations” or “loops” to generate multiple input files
- Make sure to **UPDATE** the file-path in the template file so that the input files are generated in the correct location.

______________________________________________

## 1. Write Mplus input files

```{r}

createModels(here("PYDI_enumeration_template1.txt"))

```

______________________________________________

## 2. Run models

- `recursive = TRUE` tells R to run all models within a parent folder. The recursive option is useful because when generating large batches of input files we can use the template file to create a nested set of sub-folder to organize models by type.

```{r, eval=FALSE}

runModels(here("mplus_files"), recursive = TRUE) 

```

______________________________________________

## 3. Read models

```{r}
output_enum <- readModels(here("mplus_files"), quiet = TRUE)
```

______________________________________________

## 4. Extract fit information from output files


- Any information in the ouptut or .gh5 files can be extracted and organized.
- This includes the summary statistics which are necessary to look at for choosing the number of classes. 
- The models can be sorted based on a give statistic, such as the BIC.

```{r}

showSummaryTable(output_enum, keepCols=c("Title","Parameters", "LL", "BIC","aBIC","BLRT_PValue"), sortBy="Title")

```

______________________________________________


## 5. Mplus Object lists

- Click on the Mplus object in your R environment. This is an object including nested lists. 
- This code tells R to look inside the object `output_enum` and extract the probabilities from the `.gh5` file associated with the 3-class model


```{r}

conditional_probs <- as.data.frame(output_enum[["C3_PYDI_LCA.out"]]
                           [["gh5"]]
                           [["means_and_variances_data"]]
                           [["estimated_probs"]]
                           [["values"]]
                           [seq(2, 14, 2),]) # seq("from","to","by")

conditional_probs

```

______________________________________________

## End

______________________________________________


## References

Hallquist, M. N., & Wiley, J. F. (2018). MplusAutomation: An R Package for Facilitating Large-Scale Latent Variable Analyses in Mplus. Structural equation modeling: a multidisciplinary journal, 25(4), 621-638.

Muthén, L. K., & Muthén, B. O. (2002). [\textcolor{blue}{How to use a Monte Carlo study to decide on sample size and determine power.}](https://www.tandfonline.com/doi/pdf/10.1207/S15328007SEM0904_8) Structural equation modeling, 9(4), 599-620.

R Core Team (2017). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL http://www.R-project.org/

Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686




